---
title: "Project 1: Roulette"
output: html_notebook
---

[Roulette](https://en.wikipedia.org/wiki/Roulette#Bet_odds_table){target="blank"} is an iconic game of chance, where gamblers bet on numbers selected by a ball rolling into a slot of a rotating wheel. There are various styles of wheels, but in the US the most common is the "double-zero wheel". 

```{r echo=F, fig.width=5, fig.height=5}
nums = "0-28-9-26-30-11-7-20-32-17-5-22-34-15-3-24-36-13-1-00-27-10-25-29-12-8-19-31-18-6-21-33-16-4-23-35-14-2"
strings = unlist(strsplit(nums,"-"))
ints = unlist(lapply(strings,strtoi))
N = length(strings)
theta = seq(0,2*pi,length.out=1000)
plot(sin(theta),cos(theta),type="l",ann=F,axes=F)
for(i in 1:N){
  theta = seq((i-1)/N,i/N,length.out=100)*2*pi
  x = ints[i]
  if(x==0){
    col = "green"
  } else if(i%%2==1) {
    col = "red"
  } else {
    col = "black"
  }
  polygon(c(0,sin(theta)),c(0,cos(theta)),col=col)
}
for(i in 1:N){
  lines(c(0,sin(i/N*2*pi)),c(0,cos(i/N*2*pi)),col="gray")
  itex = i-0.5
  text(0.93*sin(itex/N*2*pi),0.93*cos(itex/N*2*pi),strings[i],srt=-itex/N*360,col="white",cex=1)
}
theta = seq(0,2*pi,length.out=1000)
polygon(0.7*sin(theta),0.7*cos(theta),col="white")
```

The double-zero wheel wheel has `r N` slots. A gambler can choose [various categories](https://en.wikipedia.org/wiki/Roulette#Bet_odds_table){target="blank"} to bet on. In this paper, we will focus on two kinds of bets: straight-up and black. On a straight-up bet, the gambler chooses a specific number. On a black bet, the gambler bets the ball will select a black slot. Our gambler bets $1.00 on each spin.

For each betting style, we simulate 34 spins. This choice of 34 has two reasons: it is not used elsewhere in this paper to mean something else and it nearly maximizes the probability that a straight-up gambler will be ahead that day. (35 would maximize the chance a straight-up gambler would be ahead that day, but it also is the profit on a winning spin, so a reader may confuse the two.)

### Straight-up probability distribution

When betting straight up, there is a 1/38 chance of profiting \$35 and a 37/38 chance of profiting -\$1.00. 

| Profit |  Probability  |
|:------:|:-------------:|
| -\$1   |$\frac{37}{38}$|
| \$35   |$\frac{1}{38}$ |

We calculate the population mean.
$$\mu = \sum \text{profit} \times \text{probability}$$
```{r}
profit1 = c(-1, 35)
probs1 = c(37/38, 1/38)
mu1 = sum(profit1 * probs1)
```

$$\mu_1 = `r mu1`$$
We calculate the population standard deviation.
$$\sigma = \sum (\text{profit}-\mu)^2 \times \text{probability}$$

```{r}
sigma1 = sum((profit1-mu1)^2 * probs1)
```

$$\sigma_1 = `r sigma1`$$


### Bet-on-black probability distribution

When betting on black, there is a 18/38 chance of profiting $1 and a 20/38 chance of profiting -\$1.00.

| Profit |  Probability  |
|:------:|:-------------:|
| -\$1   |$\frac{20}{38}$|
| \$1    |$\frac{18}{38}$ |

We calculate the population mean.
```{r}
profit2 = c(-1, 1)
probs2 = c(20/38, 18/38)
mu2 = sum(profit2 * probs2)
```

$$\mu_2 = `r mu2`$$
We calculate the population standard deviation.

```{r}
sigma2 = sum((profit2-mu2)^2 * probs2)
```

$$\sigma_2 = `r sigma2`$$


