---
title: "Standard Normal Probability Formulas"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Formulas written in probability notation

## Formulas in terms of left area

### To find a right area
$$P(Z>z) ~=~ 1-P(Z<z) $$

### To find a central area (assuming $z>0$)
$$P(|Z|<z)~=~ 2P(Z<z)-1$$

### To find a two-tail area (assuming $z>0$)
$$P(|Z|>z)~=~ 2-2P(Z<z)$$

### To find an arbitrary area (assuming $z_2>z_1$)
$$P(z_1<Z<z_2) ~=~ P(Z<z_2) -P(Z<z_1)$$

## Formulas solved for left area

### To find a left area from a right area

$$P(Z<z) ~=~ 1-P(Z>z) $$

### To find a left area from a central area (assuming $z>0$)

$$P(Z<z) ~=~\frac{P(|Z|<z)+1}{2}$$

### To find a left area from a two-tail area (assuming $z>0$)

$$P(Z<z) = \frac{2-P(|Z|>z)}{2} $$

# Formulas written in nonstandard notation

Let's use a shorter notation.
$$\begin{aligned}
L &= \text{left area} &= P(Z<z) \\
R &= \text{right area} &= P(Z>z) \\
C &= \text{central area} &= P(|Z|<z) \\
T &= \text{two-tail area} &= P(|Z|>z)
\end{aligned}$$

### To find a right area
$$R ~=~ 1-L $$

### To find a central area (assuming $z>0$)
$$C~=~ 2L-1$$

### To find a two-tail area (assuming $z>0$)
$$T~=~ 2-2L$$


## Formulas solved for left area

### To find a left area from a right area

$$L ~=~ 1-R $$

### To find a left area from a central area (assuming $z>0$)

$$L ~=~\frac{C+1}{2}$$

### To find a left area from a two-tail area (assuming $z>0$)

$$L = \frac{2-T}{2} $$

## Graphic representation

```{r,fig.height=3,fig.width=6}
par(mar=c(1,0,0,0))
z = 0.8
Z = seq(-3,3,0.01)
plot(Z,dnorm(Z),type="l",ann=F,axes=F)
abline(h=0)
lines(c(z,z),c(0,dnorm(z)))
zl = (-2+z)/2
text(zl,dnorm(zl)/3,"L")
zr = (2+z)/2
text(zr,dnorm(zr)/3,"R")
mtext("z",1,at=z)
```

```{r,fig.height=3,fig.width=6}
library(latex2exp)
par(mar=c(1,0,0,0))
z = 0.8
Z = seq(-3,3,0.01)
plot(Z,dnorm(Z),type="l",ann=F,axes=F)
abline(h=0)
lines(c(-z,-z),c(0,dnorm(-z)))
lines(c(z,z),c(0,dnorm(z)))
zc = 0
text(zc,dnorm(zc)/3,"C")
zt1 = -(2+z)/2
zt2 = (2+z)/2
text(c(zt1,zt2),c(dnorm(zt1),dnorm(zt2))/3,"T/2")
mtext("-z",1,at=-z)
mtext("z",1,at=z)
```

