---
title: "Binary Random Variable"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Bernoulli Random Variable

* A Bernoulli random variable is a binary variable with two possible outcomes: 0 and 1. 
* e.g. A Bernoulli random variable with $p=\frac{3}{5}$
```{r bern_three_fifths,fig.align='center',fig.height=1,fig.width=1}
X = c(0,0,1,1,1)
N = length(X)
par(mar=c(0,0,0,0))
plot(0,0,type='n',ann=F,axes=F,xlim=c(-1,1),ylim=c(-1,1))
polygon(cos(1:1000/1000*2*pi),sin(1:1000/1000*2*pi))
for(i in 1:N){
  lines(c(0,cos(i/N*2*pi)),c(0,sin(i/N*2*pi)))
  text(0.8*cos((i-0.5)/N*2*pi),0.8*sin((i-0.5)/N*2*pi),X[i])
}
arrows(-0.6,-0.1,0.6,0.1,length=0.1)
```
* e.g. A Bernoulli random variable with $p=\frac{2}{7}$
```{r bern_two_sevenths,fig.align='center',fig.height=1,fig.width=1}
X = c(0,0,0,0,0,1,1)
N = length(X)
par(mar=c(0,0,0,0))
plot(0,0,type='n',ann=F,axes=F,xlim=c(-1,1),ylim=c(-1,1))
polygon(cos(1:1000/1000*2*pi),sin(1:1000/1000*2*pi))
for(i in 1:N){
  lines(c(0,cos(i/N*2*pi)),c(0,sin(i/N*2*pi)))
  text(0.8*cos((i-0.5)/N*2*pi),0.8*sin((i-0.5)/N*2*pi),X[i])
}
arrows(-0.6,-0.1,0.6,0.1,length=0.1)
```


## Population mean and standard deviation

* e.g. A Bernoulli variable with $p=\frac{3}{5}$
```{r bern_three_fifthsB,fig.align='center',fig.height=1,fig.width=1}
X = c(0,0,1,1,1)
N = length(X)
par(mar=c(0,0,0,0))
plot(0,0,type='n',ann=F,axes=F,xlim=c(-1,1),ylim=c(-1,1))
polygon(cos(1:1000/1000*2*pi),sin(1:1000/1000*2*pi))
for(i in 1:N){
  lines(c(0,cos(i/N*2*pi)),c(0,sin(i/N*2*pi)))
  text(0.8*cos((i-0.5)/N*2*pi),0.8*sin((i-0.5)/N*2*pi),X[i])
}
arrows(-0.6,-0.1,0.6,0.1,length=0.1)
mu = mean(X)
sig = sd(X)*sqrt((N-1)/N)
```
* Mean
$$\mu = \frac{0+0+1+1+1}{5} = \frac{3}{5} = 0.6$$
* Standard Deviation
$$\sigma = \sqrt{\frac{\left(0-\frac{3}{5}\right)^2+\left(0-\frac{3}{5}\right)^2+\left(1-\frac{3}{5}\right)^2+\left(1-\frac{3}{5}\right)^2+\left(1-\frac{3}{5}\right)^2}{5}} $$

## Simplification...

$$\begin{aligned}
\sigma &= \sqrt{\frac{\left(0-\frac{3}{5}\right)^2+\left(0-\frac{3}{5}\right)^2+\left(1-\frac{3}{5}\right)^2+\left(1-\frac{3}{5}\right)^2+\left(1-\frac{3}{5}\right)^2}{5}} \\\\
&= \sqrt{\frac{\left(\frac{3}{5}\right)^2+\left(\frac{3}{5}\right)^2+\left(\frac{2}{5}\right)^2+\left(\frac{2}{5}\right)^2+\left(\frac{2}{5}\right)^2}{5}} \\\\
&= \sqrt{\frac{2\left(\frac{3}{5}\right)^2+3\left(\frac{2}{5}\right)^2}{5}} \\\\
&= \sqrt{\frac{2}{5}\left(\frac{3}{5}\right)^2+\frac{3}{5}\left(\frac{2}{5}\right)^2}
\end{aligned}$$

## Continuation of simplification...

$$\begin{aligned}
\sigma &= \sqrt{\frac{2}{5}\left(\frac{3}{5}\right)^2+\frac{3}{5}\left(\frac{2}{5}\right)^2} \\\\
&= \sqrt{\frac{2}{5} \cdot \frac{3}{5} \cdot \left(\frac{3}{5}+\frac{2}{5} \right)} \\\\
&= \sqrt{\frac{2}{5} \cdot \frac{3}{5} \cdot  1 } \\\\
&= \sqrt{\frac{2}{5} \cdot \frac{3}{5}} = `r sqrt(2/5*3/5)`
\end{aligned}$$

## Generalization 

* You can show that any Bernoulli variable has simple formulas for mean and standard deviation.
$$\mu = p $$
$$\sigma = \sqrt{p(1-p)} $$


## Sample mean and standard deviation

* If we spin this Bernoulli variable 25 times...
```{r bern_three_fifthsC,fig.align='center',fig.height=1.5,fig.width=3}
set.seed(1234)
X = c(0,0,1,1,1)
N = length(X)
par(mar=c(0,0,0,0),mfrow=c(1,2))

plot(0,0,type='n',ann=F,axes=F,xlim=c(-1,1),ylim=c(-1,1))
polygon(cos(1:1000/1000*2*pi),sin(1:1000/1000*2*pi))
for(i in 1:N){
  lines(c(0,cos(i/N*2*pi)),c(0,sin(i/N*2*pi)))
  text(0.8*cos((i-0.5)/N*2*pi),0.8*sin((i-0.5)/N*2*pi),X[i])
}
arrows(-0.6,-0.1,0.6,0.1,length=0.1)

plot(0,0,type='n',ann=F,axes=F,xlim=c(0,15),ylim=c(0,15))
x = sample(X,25,T)
n = length(x)
text(5,12,"sample:")
for(i in 1:n){
  text(4+2*(i-1)%%5, 10-2*(i-1)%/%5, x[i],cex=0.8)
}
xb = round(mean(x),3)
s = round(sd(x),3)
suc = sum(x)
fai = n-suc
```

* We calculate the sample mean...
$$\bar{x} = \frac{`r sum(x)`}{`r n`} =`r mean(x)` $$
* We calculate the sample standard deviation...
$$s = \sqrt{\frac{`r fai`\cdot(0-`r xb`)^2+`r suc`\cdot(1-`r xb`)^2}{`r n`-1}} = `r sd(x)` $$

## Sample mean and standard deviation again

* If we spin this Bernoulli variable 25 times...
```{r,fig.align='center',fig.height=1.5,fig.width=3}
X = c(0,0,1,1,1)
N = length(X)
par(mar=c(0,0,0,0),mfrow=c(1,2))

plot(0,0,type='n',ann=F,axes=F,xlim=c(-1,1),ylim=c(-1,1))
polygon(cos(1:1000/1000*2*pi),sin(1:1000/1000*2*pi))
for(i in 1:N){
  lines(c(0,cos(i/N*2*pi)),c(0,sin(i/N*2*pi)))
  text(0.8*cos((i-0.5)/N*2*pi),0.8*sin((i-0.5)/N*2*pi),X[i])
}
arrows(-0.6,-0.1,0.6,0.1,length=0.1)

plot(0,0,type='n',ann=F,axes=F,xlim=c(0,15),ylim=c(0,15))
x = sample(X,25,T)
n = length(x)
text(5,12,"sample:")
for(i in 1:n){
  text(4+2*(i-1)%%5, 10-2*(i-1)%/%5, x[i],cex=0.8)
}
xb = round(mean(x),3)
s = round(sd(x),3)
suc = sum(x)
fai = n-suc
```

* We calculate the sample mean...
$$\bar{x} = \frac{`r sum(x)`}{`r n`} =`r mean(x)` $$
* We calculate the sample standard deviation...
$$s = \sqrt{\frac{`r fai`\cdot(0-`r xb`)^2+`r suc`\cdot(1-`r xb`)^2}{`r n`-1}} = `r sd(x)` $$



## Repeat many times (10000 iterations of samples of size 25)

```{r,cache=T}
set.seed(1234)
xbars = numeric()
for(i in 1:10000){
  xbars = c(xbars,mean(sample(X,25,T)))
}
hist(xbars,breaks=seq(min(x)-1/50,max(x)+1/50,by=1/25))
```

* Estimate **the standard error** (the standard deviation of repeated sample means)

## Standard error

* The standard error is the standard deviation of repeated sample means.
* We will use $\text{SE}$ and $\sigma_{\bar{x}}$ to denote the standard error.
* We learned that standard deviation of a bell is approximately $\text{range}/6$.
$$\frac{0.9-0.3}{6} = `r (0.9-0.3)/6` $$
* We will later show that we can calculate standard error if we know $\sigma$.
$$\sigma_{\bar{x}} = \frac{\sigma}{\sqrt{n}} $$
* Notice, in our example, we can calculate the standard error using this formula.
$$\sigma_{\bar{x}} = \frac{`r sig`}{\sqrt{`r n`}} = `r sig/sqrt(n)` $$


## Interval of typical sample means

We will define the **interval of typical sample means** as:
$$\text{interval of typical means} = \left(\mu-\frac{2\sigma}{\sqrt{n}},\,\mu+\frac{2\sigma}{\sqrt{n}}\right)$$

* We expect about 95% of sample means to be in the interval of typical sample means.
* This is more true when $n$ is larger or when the random variable is nearly normal.

## Back to example.

$$\mu=p=\frac{3}{5}=0.6$$

$$\sigma = \sqrt{\frac{3}{5}\cdot \frac{2}{5}} = `r sqrt(2/5*3/5)` $$

$$n=25$$

$$\begin{aligned}
\text{interval of typical means} &= \left(\mu-\frac{2\sigma}{\sqrt{n}},\,\mu+\frac{2\sigma}{\sqrt{n}}\right) \\
&= \left(`r mu`-2\cdot\frac{`r sig`}{\sqrt{`r n`}},\,`r mu`+2\cdot\frac{`r sig`}{\sqrt{`r n`}}\right) \\
&= \left(`r round(mu-2*sig/sqrt(n),3)`,\,`r round(mu+2*sig/sqrt(n),3)`\right)
\end{aligned}$$

## What percent of sample means were between 0.404 and 0.796?

```{r, fig.height=5, fig.width=9}
myh = hist(xbars,labels=T,breaks=seq(min(x)-1/50,max(x)+1/50,by=1/25),cex=0.4,ylim=c(0,1600),xlim=c(0.3,0.9))
typicals = myh$counts[(myh$mids > 0.4) & (myh$mids < 0.8)]
str1 = paste(typicals,collapse = "+")
```

\small
$$\frac{`r str1`}{`r length(xbars)`} $$

$$`r sum(typicals)/length(xbars)` =  `r sum(typicals)/length(xbars)*100` \% $$

\normalsize

## Derivation of standard error

$$SD\left(\frac{X_1+X_2+X_3+\cdots+X_{24}+X_{25}}{25} \right) $$
$$\sqrt{\frac{SD(X_1)^2}{25^2}+\frac{SD(X_2)^2}{25^2}+\cdots+\frac{SD(X_{25})^2}{25^2}}$$

$$\sqrt{\frac{25\cdot SD(X)^2}{25^2}}$$

$$\frac{SD(X)}{5}$$