```{r data generation, echo = FALSE, results = "hide"}
ell = runif(1,0.01,0.99)
z = round(qnorm(ell),2)
ell = round(pnorm(z),4)
sol = 1-ell
```

Question
========
Determine the probability that the standard normal variable is more than `r z`. In other words, evaluate $P(Z > `r z`)$.


Solution
========
First, you need to identify that we are looking for a right area. This is because large values of $Z$ satisfy $Z>`r z`$.

By using the z-table, we find the left area (even though we eventually want the right area). 

$$P(Z<`r z`)=`r ell`$$

We use the rule of compliments to determine the right area.

$$\begin{aligned}
P(Z > `r z`) &= 1-P(Z < `r z`) \\
&= 1-`r ell` \\
&= `r sol` 
\end{aligned}$$

Method 2: You might also recognize that the normal distribution is symmetric. Thus,

$$\begin{aligned}
P(Z > `r z`) &= P(Z < `r -z`) \\
&= `r sol` 
\end{aligned}$$

Method 3: It often helps to draw a picture. 

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=z],3,z),c(dnorm(zs[zs>=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
midl = (-3+z)/2
midr = (3+z)/2
text(-2,0.3,paste("left area =",round(pnorm(z),4)))
arrows(-2,0.28,midl,dnorm(midl)/2)
text(2,0.3,paste("right area =",round(1-pnorm(z),4)))
arrows(2,0.28,midr,dnorm(midr)/2)
axis(1,at=c(-3,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\


Meta-information
================
extype: num
exsolution: `r sol`
exname: Get rightarea from Z
extol: 0.0002
