```{r data generation, echo = FALSE, results = "hide"}
ell = runif(1,0.65,0.98)
z = round(qnorm(ell),2)
ell = round(pnorm(z),4)
la = round(pnorm(-z),4)
sol = 2*la
```

Question
========
Determine the probability that the **absolute** standard normal variable is more than `r z`. In other words, evaluate $P\left(|Z| > `r z`\right)$.


Solution
========
First, you need to identify that we are looking for a two-tail area (the sum of the left and right tails). This is because $z$-scores far from 0 satisfy $|Z|>`r z`$, while $z$-scores near 0 do not satisfy the inequality.

Start with a sketch.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),c(dnorm(zs[zs<=-z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),c(dnorm(zs[zs>=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\

Remember, the entire area is always 1. Our z-table only provides left areas. 

**Method 1:** First find the area of the left tail.
$$P(Z<`r -z`) =  `r la`$$
Recognize normal distributions are symmetric, so we also know the area of the right tail.
$$P(Z>`r z`) = `r la` $$
The two areas add to our desired two-tail area.
$$\begin{aligned}
P\left(|Z|>`r z`\right) &= P(Z<`r -z`)+P(Z>`r z`) \\
&= `r la` + `r la` \\
&= `r sol`
\end{aligned}$$

This technique can be summarized with the following formula:
$$P\left(|Z|>z\right) = 2\cdot P(Z<-z)$$
assuming $z>0$.

So, in our case (when $z=`r z`$),
$$\begin{aligned}
P\left(|Z|<`r z`\right) &= 2\cdot P(Z<`r -z`) \\
&= 2\cdot`r la` \\
&= `r sol`
\end{aligned}$$

This method is shown graphically:

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),c(dnorm(zs[zs<=-z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),c(dnorm(zs[zs>=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/2,0.01),la)
text(xr,max(dnorm(xr)/2,0.01),la)
```
\

Notice we need to add both tails.

**Method 2:** You can achieve the same result by using the following formula:
$$P\left(|Z|>z\right) = 2\cdot\left[1-P(Z<z) \right] $$
So,
$$\begin{aligned}
P\left(|Z|>`r z`\right) &= 2\cdot\left[1-P(Z<`r z`) \right] \\
&= 2\cdot\left[1-`r ell`\right] \\
&= 2\cdot`r la`\\
&= `r sol`
\end{aligned}$$
```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),c(dnorm(zs[zs<=-z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),c(dnorm(zs[zs>=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(-3,z,z),c(0.42,0.42,0.4))
text((-3+z)/2,0.45,ell)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/2,0.01),la)
text(xr,max(dnorm(xr)/2,0.01),la)
```
\



Meta-information
================
extype: num
exsolution: `r sol`
exname: Get two-tail area from Z
extol: 0.0002
