```{r data generation, echo = FALSE, results = "hide"}
while(T)
{
  ell = round(runif(1,0.01,0.99),2)
  z = qnorm(ell)
  zlow = floor(z*100)/100
  zhigh = ceiling(z*100)/100
  if(round(z,4) != round(z,2)) {break}
}
```

Question
========
Determine $z$ such that $P(Z<z)=`r ell`$. In other words, what $z$-score is greater than $`r ell*100`$% of standard normal values? (Answers within 0.01 from the correct value will be marked correct.)


Solution
========
Start with a sketch.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=z],z,-3),
        c(dnorm(zs[zs<=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text((-3+z)/2,max(dnorm((-3+z)/2)/3,0.02),ell)
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\


```{r, echo=F}
z2 = seq(zlow-0.02,zhigh+0.02,0.01)
```

You should go to your $z$-table and find the $z$-score with the left area closest to `r ell`.

| $z$ | $P(Z<z)$ |
|:---:|:---:|
| `r z2[1]` | `r format(round(pnorm(z2[1]),4),4)` |
| `r z2[2]` | `r format(round(pnorm(z2[2]),4),4)` |
| **`r z2[3]`** | **`r format(round(pnorm(z2[3]),4),4)`** |
| **`r z2[4]`** | **`r format(round(pnorm(z2[4]),4),4)`** |
| `r z2[5]` | `r format(round(pnorm(z2[5]),4),4)` |
| `r z2[6]` | `r format(round(pnorm(z2[6]),4),4)` |

It turns out the exact answer is `r z`, which could be found by using an inverse normal function. Like InvNorm(`r ell`) on a TI calculator, Norm.Inv(`r ell`) on a spreadsheet, or qnorm(`r ell`) in R.

But, because we are using the $z$-table, I will accept either `r zlow` or `r zhigh`. (Or, really anything within 0.011 of `r z`.)



Meta-information
================
extype: num
exsolution: `r z`
exname: Get z from left area
extol: 0.011
