```{r data generation, echo = FALSE, results = "hide"}
## DATA GENERATION

shapes = c("tight","uniform","bimodal")

b = c(3,1,0.4)
r = sort(sample(c(10,20,30,40),3,replace=T))
m = sample(c(20,30,40,50,60),3)
den = c(6,4,2)

x1 = rbeta(10000,b[1],b[1])*r[1]+m[1]-r[1]/2
x2 = rbeta(10000,b[2],b[2])*r[2]+m[2]-r[2]/2
x3 = rbeta(10000,b[3],b[3])*r[3]+m[3]-r[3]/2

xs = c(x1,x2,x3)

ord = sample(0:2)
let = c("F","G","H")

xx1 = xs[(ord[1]*10000+1):(ord[1]*10000+10000)]
xx2 = xs[(ord[2]*10000+1):(ord[2]*10000+10000)]
xx3 = xs[(ord[3]*10000+1):(ord[3]*10000+10000)]
```

Question
========
We can estimate the mean of a **symmetric** distribution.
$$\bar{x} \approx \frac{\max(x)+\min(x)}{2} $$

We can **roughly** estimate the standard deviation of certain distributions.

| Shape | SD estimate |
|:-----:|:-----------:|
| bell  | $\text{range}/6$ |
| uniform | $\text{range}/4$ |
| bimodal | $\text{range}/2$ |


Three random variables (F, G, and H) were measured 10000 times each. The resulting histograms show the three distributions.

\
```{r, showitnow, echo=FALSE, fig.height = 5, fig.width=7, fig.path=paste("hists",sample(1:9999,1),sep="")}
par(mar=c(3,0,2,0))
layout(matrix(c(1,2,3), 3, 1, byrow = TRUE))
hist(xx1,breaks=(0:80),ann=F,yaxt="n",xaxt="n")
mtext("Random Variable F",side=3)
axis(1,at=0:16*5)
hist(xx2,breaks=(0:80),ann=F,yaxt="n",xaxt="n")
mtext("Random Variable G",side=3)
axis(1,at=0:16*5)
hist(xx3,breaks=(0:80),ann=F,yaxt="n",xaxt="n")
mtext("Random Variable H",side=3)
axis(1,at=0:16*5)
```
\

Answerlist
--------
* Estimate the mean of F.
* Estimate the mean of G.
* Estimate the mean of H.
* Estimate the standard deviation of F.
* Estimate the standard deviation of G.
* Estimate the standard deviation of H.


Solution
========

Answerlist
--------
* `r m[ord[1]+1]`
* `r m[ord[2]+1]`
* `r m[ord[3]+1]`
* `r r[ord[1]+1]/den[ord[1]+1]`
* `r r[ord[2]+1]/den[ord[2]+1]`
* `r r[ord[3]+1]/den[ord[3]+1]`



Meta-information
============
extype: string
exsolution: yup
exname: estimate mean and sd
extol: 0.01
