```{r,data,generation,echo=FALSE,results="hide"}
type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
n = sample(7:15,1)
beta1 = c(1,5,0.2,0.5,2)[type]
beta2 = c(1,5,0.2,2,0.5)[type]
bottom = sample(seq(20,70,5),1)
width = sample(seq(5,30,5),1)
i = 1:n
x = round(rbeta(n,beta1,beta2)*width+bottom,2)
```


Question
========

A sample was gathered. 

$$\mathbf{x} = `r x` $$


```{r echo=F}
write.csv(data.frame(i,x),file="get_median.csv", quote = FALSE, row.names = FALSE)
```

You can download the data as a [CSV file](get_median.csv).

Determine the sample [median](https://en.wikipedia.org/wiki/Median){target="blank"}. Please enter an exact answer.



Solution
========

To determine the median by hand, you first sort the sample.
$$\text{sort}(\mathbf{x})=`r sort(x)`$$
If the sample size is odd, just find the number in the middle number (the $\frac{n+1}{2}$th value). If the sample size is even, determine the mean of the middle two numbers (the $\frac{n}{2}$th and $\left(\frac{n}{2}+1\right)$th values).


```{r echo=F}
sol = data.frame(i,x)
sol = rbind(sol,c("",""),c("","median(x)"),c("",paste0("=median(B2:B",n+1,")",collapse="")))
write.csv(sol,"median_solution.csv",row.names=F)
```

In a spreadsheet, you can use the [`MEDIAN` function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_MEDIAN_function){target="blank"}. You can see a [solution spreadsheet here](median_solution.csv).

In R, you can use the [`MEDIAN` function](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/median){target="blank"}.

```{r}
# First, get the csv into the working directory... then...
data = read.csv("get_median.csv")
x = data$x
median(x)
```



Meta-information
================
extype: num
exsolution: `r median(x)`
exname: get_median
extol: 0
exextra[numwidth]: 999.999