```{r data generation, echo = FALSE, results = "hide"}
while(T){
  n = sample((6:17)^2,1)
  p = round(runif(1,0.05,0.95),2)
  sig = sqrt(p*(1-p)/n)
  if(n*p>12 & n*(1-p)>12){break}
}

ell = round(0.5+sample(c(-1,1),1)*runif(1,0.04,0.48),3)
z = round(qnorm(ell),2)
phat = round(p + z*sig,2)
z = round((phat-p)/sig,2)
P = round(pnorm(z),4)
```

Question
========

Let random variable $W$ have the probability distribution shown below.

\begin{center}
\begin{tabular}{|c|c|} \hline
$w$ & $P(w)$ \\ \hline
0 & `r 1-p` \\
1 & `r p` \\ \hline
\end{tabular}
\end{center}

Let random variable $\hat{p}$ (sample proportion) represent the average of $n=`r n`$ instances of $W$. 

Answerlist
--------

* Determine the mean and standard deviation of the $\hat{p}$. \newline\vfill
* Determine $P(\hat{p} < `r phat`)$. Do NOT use a continuity correction. **Draw a sketch** \newline\vfill


Solution
========

\setkeys{Gin}{width=.5\textwidth}

Answerlist
--------
* We can recognize $W$ is a Bernoulli variable with $p=`r p`$ and $q=`r 1-p`$. Thus, $$\mu_{\tiny w} = p = `r p`$$ and $$\sigma_{\tiny w} = \sqrt{pq} = \sqrt{(`r p`)(`r 1-p`)} = `r sqrt(p*(1-p))`$$  \newline
  Then, we use the central limit of average formulas: $$\mu_{\tiny \hat{p}} = \mu_{\tiny w} = `r p`$$ and $$\sigma_{\tiny \hat{p}} = \frac{\sigma_{\tiny w}}{\sqrt{n}} = \frac{`r sqrt(p*(1-p))`}{\sqrt{`r n`}} = `r sig`$$  \newline
  But, if we recognized $\hat{p}$ follows the formulas of a **$\hat{p}$ sampling distribution**:
  $${\text{\Large $\mu$}}_{\tiny \hat{p}} = p $$
  $${\text{\Large $\sigma$}}_{\tiny \hat{p}} = \frac{\sqrt{pq}}{\sqrt{n}} = \sqrt{\frac{pq}{n}} $$
  we could have just used those instead.
* $P(\hat{p} < `r phat`) ~=~ P(Z < `r z`) ~=~ `r P`$
\newline
```{r no7,echo=F,fig.height=2,fig.width=7,fig.cap=""}
library(latex2exp)
par(mar=c(4,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
Z = -350:350/100
lines(Z,dnorm(Z))
abline(h = 0)
axis(1,-3:3,labels=round((-3:3)*sig+p,3),line=2, tick = FALSE)
mtext(TeX("$\\hat{p}=$"),1,3,adj = 0)
polygon(c(Z[Z<=z],z,-3.5),c(dnorm(Z[Z<=z]),0,0),col=rgb(0.4,0.4,0.4,0.4))
```
\setkeys{Gin}{width=\textwidth}


Meta-information
============
extype: string
exsolution: yo
exname: marbles
extol: 0.01
