```{r data generation, echo = FALSE, results = "hide"}
mu = sample(46:120,1)
sig = sample(3:(mu%/%3),1)
p1 = round(runif(1,0.02,0.98),3)
p2 = round(runif(1,0.02,0.98),3)
p3 = round(runif(1,0.52,0.98),3)
p4 = round(runif(1,0.52,0.98),3)
p = c(p1,p2,p3,p4)

z = round(qnorm(p),1)
x = signif(mu+z*sig,4)
z = round((x-mu)/sig,2)
p = round(pnorm(z),4)

P1 = p[1]
P2 = 1-p[2]
P3 = p[3]*2-1
P4 = (1-p[4])*2
P = c(P1,P2,P3,P4)

a = x-mu
```

Question
========

Let $X$ be normally distributed with mean `r mu` and standard deviation `r sig`. Please calculate the probabilities shown below and also shade a corresponding region under the density curve.

\setkeys{Gin}{width=.6\textwidth}

Answerlist
--------

* $P(X<`r x[1]`)$  \newline \vfill \hfill
```{r no1,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
```
* $P(X>`r x[2]`)$  \newline \vfill \hfill
```{r no2,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
``` 
* $P(`r signif(mu-z[3]*sig,3)` <X< `r x[3]`)$  \newline \vfill \hfill
```{r no3,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
```

\setkeys{Gin}{width=1\textwidth}

Solution
========

\setkeys{Gin}{width=.6\textwidth}

Answerlist
--------
* $P(X<`r x[1]`) ~=~ P(Z<`r z[1]`) ~=~ \fbox{`r P1`}$ \newline
```{r no5,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
Z = -350:350/100
lines(Z,dnorm(Z))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
polygon(c(Z[Z<=z[1]],z[1],-3.5),c(dnorm(Z[Z<=z[1]]),0,0),col=rgb(0.4,0.4,0.4,0.4))
```
* $P(X>`r x[2]`) ~=~ P(Z>`r z[2]`) ~=~ \fbox{`r P2`}$ \newline
```{r no6,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
Z = -350:350/100
lines(Z,dnorm(Z))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
polygon(c(Z[Z>=z[2]],3.5,z[2]),c(dnorm(Z[Z>=z[2]]),0,0),col=rgb(0.4,0.4,0.4,0.4))
```
* $P(`r signif(mu-z[3]*sig,3)` <X< `r x[3]`) ~=~ P(`r -z[3]`<Z<`r z[3]`) ~=~ \fbox{`r P3`}$ \newline
```{r no7,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
Z = -350:350/100
lines(Z,dnorm(Z))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
polygon(c(Z[Z>=-z[3] & Z<=z[3]],z[3],-z[3]),c(dnorm(Z[Z>=-z[3] & Z<=z[3]]),0,0),col=rgb(0.4,0.4,0.4,0.4))
```

\setkeys{Gin}{width=\textwidth}



Meta-information
============
extype: string
exsolution: yo
exname: marbles
extol: 0.01
