---
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\thispagestyle{fancy}

\chead{PRACTICE}
\rhead{Formulas}


### Normal standardization
$$\begin{aligned}
z &= \frac{x-\mu}{\sigma} &&&&& x &= \mu + z\sigma
\end{aligned}$$

### Central Limit Theorem

If:

- $W$ is "any" random variable with mean = $\mu_{\tiny w}$ and standard deviation = $\sigma_{\tiny w}$.
- Random variable $X$ is **sum** of $n$ instances of $W$. $$X = W_1+W_2+W_3+\cdots+W_n $$
- Random variable $Y$ is **average** of $n$ instances of $W$. $$Y = \frac{W_1+W_2+W_3+\cdots+W_n}{n} $$

Then:

- The following formulas are exactly true:
$$\begin{aligned} 
\mu_{\tiny x} &= n \mu_{\tiny w} &&&&& \mu_{\tiny y} &= \mu_{\tiny w} \\
\sigma_{\tiny x} &= \sigma_{\tiny w} \sqrt{n} &&&&& \sigma_{\tiny y} &= \frac{\sigma_{\tiny w}}{\sqrt{n}} 
\end{aligned}$$
- $X$ and $Y$ are approximately normal (if $n\ge 30$)
- $X$ and $Y$ are exactly normal if $W$ is normal

### Special case of central limit theorem: Bernoulli, Binomial, and $\hat{p}$ sampling

If:

- $W$ is a Bernoulli random variable:
    \newline\begin{center}\begin{tabular}{|c|c|}\hline
    $w$ & $P(w)$ \\ \hline
    0 & $q$ \\
    1 & $p$ \\ \hline
    \end{tabular}\end{center}
- $X$ is sum of $n$ instances of $W$ ($X$ is binomial)
- $\hat{p}$ is average of $n$ instances of $W$ (proportion sampling)

Then:

- The following are exactly true:
$$\begin{aligned} 
\mu_{\tiny w}&=p &&&&&
\mu_{\tiny x} &= np &&&&& 
\mu_{\tiny \hat{p}} &= p  \\
\sigma_{\tiny w}&=\sqrt{pq} &&&&&
\sigma_{\tiny x} &= \sqrt{pq}\sqrt{n} &&&&& 
\sigma_{\tiny \hat{p}} &= \frac{\sqrt{pq}}{\sqrt{n}} 
\end{aligned}$$
- $X$ and $\hat{p}$ are approximately normal (if $np\ge 10$ and $nq\ge 10$)

\newpage


## Question 1

\rhead{}

```{r q1dg, echo = FALSE, results = "hide"}
set.seed(1234)
types = c("A","B","C","D")
mus = c(80,130,95,115)
sigs = c(8,10,15,5)
zs = c(1.5,1.2,-0.9,-1.2)
xs = signif(mus+zs*sigs,4)
zs = round((xs-mus)/sigs,2)
ps = round(pnorm(zs),4)
```

A farm produces 4 types of fruit: $A$, $B$, $C$, and $D$. The fruits' masses follow normal distributions, with parameters dependent on the type of fruit.

\begin{center}
\begin{tabular}{|c|c|c|}\hline
Type of fruit & Mean mass (g) & Standard deviation of mass (g) \\ \hline
$A$ & `r mus[1]` & `r sigs[1]` \\
$B$ & `r mus[2]` & `r sigs[2]` \\
$C$ & `r mus[3]` & `r sigs[3]` \\
$D$ & `r mus[4]` & `r sigs[4]` \\ \hline
\end{tabular}
\end{center}

One specimen of each type is weighed. The results are shown below.

\begin{center}
\begin{tabular}{|c|c|}\hline
Type of fruit & Mass of specimen (g) \\ \hline
$A$ & `r xs[1]` \\
$B$ & `r xs[2]` \\
$C$ & `r xs[3]` \\
$D$ & `r xs[4]` \\ \hline
\end{tabular}
\end{center}

a. Determine a $z$-score for each specimen. \vfill \vfill
b. Which specimen was most unusually large? \vfill
c. Which specimen was most unusually small? \vfill

\newpage




## Question 2

```{r q2dg, echo = FALSE, results = "hide"}
set.seed(147)
mu = 77
sig = 13
```

A normal random variable $X$ has a mean $\mu=`r mu`$ and standard deviation $\sigma=`r sig`$. Please label the density curve with:

a. The appropriate values of $x$.
b. The areas of the sections.

\vfill
```{r normal,echo=F,fig.width=7,fig.height=5,fig.cap=""}
library(latex2exp)
zs = (-35):35/10
par(mar=c(6,1,1,1))
plot(0,0,type="n",ann=F,axes=F,xlim=c(-3.5,3.5),ylim=c(0,0.42))
lines(zs,dnorm(zs))
zz = (-3):3
axis(at=zz,side=1)
mtext(TeX("$z=$"),side=1,line=1,adj=0)
for(z in -2:2){
  lines(c(z,z),c(0,dnorm(z)))
}
axis(at=zz,side=1,line=3,labels = rep("___",length(zz)),cex.axis=2,tick = FALSE)
mtext(TeX("$x=$"),side = 1,line=4,adj=0)
arrows(-3,0.2,-2.5,dnorm(-2.5)/3)
arrows(3,0.2,2.5,dnorm(2.5)/3)
arrows(-2,0.3,-1.5,dnorm(-1.5)/2)
arrows(2,0.3,1.5,dnorm(1.5)/2)
arrows(-1,0.4,-0.5,dnorm(-0.5)/2)
arrows(1,0.4,0.5,dnorm(0.5)/2)
xxx = c(-3,-2,-1,1,2,3)
yyy = c(0.21,0.31,0.41,0.41,0.31,0.21)
for(i in 1:length(xxx)){
  text(xxx[i],yyy[i],"___",cex=2)
}
```
\vfill

## Question 3

```{r q3dg, echo = FALSE, results = "hide"}
mu = 25
sig = 5

z = c(-0.8,-0.4,1.6)
x = signif(mu+z*sig,4)
z = round((x-mu)/sig,2)
p = round(pnorm(z),4)

P1 = p[1]
P2 = 1-p[2]
P3 = p[3]*2-1
P4 = (1-p[4])*2
P = c(P1,P2,P3,P4)

```

Let $X$ be normally distributed with mean `r mu` and standard deviation `r sig`. Please calculate the probabilities shown below and also shade a corresponding region under the density curve.

\setkeys{Gin}{width=.7\textwidth}


a. $P(X<`r x[1]`)$  \newline \vfill \hfill
```{r no1,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
```
\vfill
b. $P(X>`r x[2]`)$  \newline \vfill \hfill
```{r no2,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
``` 
\vfill
c. $P(`r signif(mu-z[3]*sig,3)` <X< `r x[3]`)$  \newline \vfill \hfill
```{r no3,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
```

\setkeys{Gin}{width=1\textwidth}
\newpage

## Question 4

Let random variable $W$ have a **mean of $\mu_{\tiny w} = 10.50$** and a **standard deviation of $\sigma_{\tiny w}= 5.77$**. 

Let $X$ represent the sum of $n=81$ instances of $W$, and let $Y$ represent the average of $n=81$ instances of $W$.

a. Determine $\mu_{\tiny x}$\vfill
b. Determine $\sigma_{\tiny x}$\vfill
c. What is the probability that $X$ is between 773 and 928? *Do NOT use a continuity correction.*\vfill\vfill\vfill
d. Determine $\mu_{\tiny y}$\vfill
e. Determine $\sigma_{\tiny y}$\vfill
f. What is the probability that $Y$ is between 10.2 and 10.8? *Do NOT use a continuity correction.*\vfill\vfill\vfill

\newpage

## Question 5

An unfair coin has a $p=0.63$ chance of landing tails. When $n=100$ of these unfair coins are flipped, what is the probability of getting at least 55 but at most 71 tails? Please use a **normal approximation** with a **continuity correction**.

\newpage

## Question 6

About 8% of men are color blind ($p=0.08$). If you gather a simple random sample of $n=121$ men, what is the probability that the sample proportion ($\hat{p}$) is between 0.03 and 0.07? Please use a **normal approximation**, but *do NOT use a continuity correction*.

\newpage

## Extra Credit

1. Let random variable $X$ be normally distributed with mean $\mu=85$ and standard deviation $\sigma=10$. Determine $a$ such that $P(|X-85|<a) = 0.8064$.
\vfill
2. A population has a proportion $p=0.7$. The sample proportion is $\hat{p}$. Determine $n$ such that $P(0.68<\hat{p}<0.72) = 0.9544$. In other words, determine the necessary sample size such that the sample proportion is between 0.68 and 0.72 in 95.44% of random samples of that size.
\vfill

