---
output:
  pdf_document: default
  html_document: default
---
```{r data generation, echo = FALSE, results = "hide"}
## DATA GENERATION
library(latex2exp)
options(scipen=999)
while(T){
  scalex = 10^sample(0:2,1)
  scaley = 10^sample(0:2,1)
  n = sample(6:9,1)
  x = signif(runif(n,1,10)*scalex,2)
  a = runif(1,1,10)*scaley
  a2 = runif(1,1,10)*scaley
  b = (a2-a)/scalex/10
  y = signif(a+b*x+rnorm(n,0,runif(1,0.3,3)*scaley),2)
  
  n5 = length(x)
  xbar5 = signif(mean(x),4)
  ybar5 = signif(mean(y),4)
  sx5 = signif(sd(x),4)
  sy5 = signif(sd(y),4)
  r5 = signif((sum(x*y)-n5*xbar5*ybar5)/(n5-1)/sx5/sy5,3)
  b5 = signif(r5*sy5/sx5,3)
  a5 = signif(mean(y)-b5*mean(x),3)
  a6 = a5+b5*(10*scalex)
  
  if(sum(y<0)==0 & sum(y>10*scaley)==0 & abs(a5-a6)>2*scaley  & a5>0 & a5<10*scaley & a6>0 & a6<10*scaley & scaley*scalex < 10^4 ){break}
}
s = '\\begin{tabular}{|c|c|c|}\\hline\n $x$ & $y$ & $\\phantom{xxxx}xy\\phantom{xxxx}$ \\\\ \\hline \n'
for(i in 1:length(x)){
  s = paste(s,x[i]," & ",y[i]," & \\\\ \n",sep="")
}
s = paste(s,'\\hline \n $\\sum x = \\phantom{xxxxxxxx}$ & $\\sum y = \\phantom{xxxxxxxx}$ & $\\sum xy = \\phantom{xxxxxxxx}$ \\\\ \\hline \n $\\bar{x}= \\phantom{xxxxx}$ & $\\bar{y}= \\phantom{xxxxx}$ & \\\\ \\hline \n')
s = paste(s,'$s_x= \\phantom{xxxxx}$ & $s_y= \\phantom{xxxxx}$ & \\\\')
s = paste(s,'\\hline \n \\end{tabular}')
```

Question
========

(20 points) [Note: this question uses 2 pages.]

You have collected the following data:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\large
`r s`
\end{center}

Answerlist
--------
* Complete the table.
* Calculate the correlation coefficient ($r$) using the formula below. 
$$r = \frac{\sum xy - n \bar{x}\bar{y}}{(n-1)s_x s_y} $$
\vfill
* The least-squares regression line will be represented as $y = a + bx$. Determine the parameters ($b$ and $a$) using the formulas below.
$$b = r \frac{s_y}{s_x} $$
$$a = \bar{y}-b \bar{x} $$
\vfill
* Write the equation of the regression line (using the calculated values of $a$ and $b$.)
\vfill
\newpage
* Please plot the data and a corresponding regression line.

\vfill \setkeys{Gin}{width=\linewidth}
  
\ \hspace{-50pt}
```{r makegridF,fig.width=5,fig.height=5,echo=F}
possmax = c(1,10,100,1000,10000)
myl = min(possmax[possmax>max(max(y),a)])
par(mar=c(4,4,1,1))
plot(x,y,type="n",xlim=c(0,1)*10*scalex,ylim=c(0,myl))
for(i in seq(0,1,0.01)){
  lines(c(i,i)*10*scalex,c(0,1)*myl,lwd=0.3,col=rgb(0.6,0.6,0.6))
  lines(c(0,1)*10*scalex,c(i,i)*myl,lwd=0.3,col=rgb(0.6,0.6,0.6))
}
for(i in seq(0,1,0.1)){
  lines(c(i,i)*10*scalex,c(0,1)*myl,lwd=0.7)
  lines(c(0,1)*10*scalex,c(i,i)*myl,lwd=0.7)
}
```
\

\vfill

Solution
========

Remember the formula for the correlation coefficient.

$$r = \frac{\sum x_i y_i - n\bar{x}\bar{y}}{(n-1)s_x s_y}$$

We calculate the necessary values.

```{r cakc, echo=F}
s2 = '\\begin{tabular}{|c|c|c|}\\hline\n $x$ & $y$ & $xy$ \\\\ \\hline \n'
for(i in 1:length(x)){
  s2 = paste(s2,x[i]," & ",y[i]," & ",x[i]*y[i]," \\\\ \n",sep="")
}
s2 = paste(s2,'\\hline \\hline \n $\\sum x = ',sum(x),'$ & $\\sum y = ',sum(y),'$ & $\\sum x_i y_i = ',sum(x*y),'$ \\\\ \\hline \n',sep="")
s2 = paste(s2,'\\hline \n $\\bar{x} = ',signif(mean(x),4),'$ & $\\bar{ y} = ',signif(mean(y),4),'$ &  \\\\ \\hline \n',sep="")
s2 = paste(s2,'\\hline \n $s_x = ',signif(sd(x),4),'$ & $s_y = ',signif(sd(y),4),'$ &  \\\\ \\hline \n',sep="")
s2 = paste(s2,' \n \\end{tabular}')

n = length(x)
xbar = signif(mean(x),4)
ybar = signif(mean(y),4)
sx = signif(sd(x),4)
sy = signif(sd(y),4)
r = signif((sum(x*y)-n*xbar*ybar)/(n-1)/sx/sy,3)
b = signif(r*sy/sx,3)
a = signif(mean(y)-b*mean(x),3)
```

`r s2`

$$r = \frac{`r sum(x*y)` - (`r n`)(`r xbar`)(`r ybar`)}{(`r n`-1)(`r sx`)(`r sy`)} = `r r` $$
If you didn't round any of the steps up to here, you'd get an exact value which is pretty close to our value.
$$r_{\text{exact}} = `r cor(x,y)` $$

The regression line has the form
$$y = a + bx $$
So, $a$ is the $y$-intercept and $b$ is the slope. We have formulas to determine them:
$$b ~=~ r \frac{s_y}{s_x} ~=~  `r r` \cdot \frac{`r sy`}{`r sx`} ~=~ `r b`$$
$$a ~=~ \bar{y}-b\bar{x} ~=~ `r signif(mean(y),3)`-(`r b`)(`r signif(mean(x),3)`) ~=~ `r a` $$ 
Our regression line:
$$y = `r a`+(`r b`)x $$

Make a plot. \setkeys{Gin}{width=0.6\linewidth}

\
```{r makeansF,fig.width=5,fig.height=5,echo=F}
possmax = c(1,10,100,1000,10000)
myl = min(possmax[possmax>max(max(y),a)])
plot(x,y,xlim=c(0,1)*10*scalex,ylim=c(0,myl))
for(i in seq(0,1,0.01)){
  lines(c(i,i)*10*scalex,c(0,1)*myl,lwd=0.3,col=rgb(0.6,0.6,0.6))
  lines(c(0,1)*10*scalex,c(i,i)*myl,lwd=0.3,col=rgb(0.6,0.6,0.6))
}
for(i in seq(0,1,0.1)){
  lines(c(i,i)*10*scalex,c(0,1)*myl,lwd=0.7)
  lines(c(0,1)*10*scalex,c(i,i)*myl,lwd=0.7)
}
abline(a,b)
```
\



Meta-information
============
extype: num
exsolution: 0
exname: regression
extol: 0.01
