---
title: "Archery"
author: "Chad Worley"
date: "8/17/2020"
output:
  html_notebook:
    code_folding: hide
---

## Archery Scoring

A bullseye is worth 10 points, and the concentric circles are worth 9, 8, 7,..., 2, 1; a miss awards 0 points.

```{r bullseye,fig.width=8,fig.height=8}
makebullseye = function(){
  x = cos(2*seq(0,pi,length.out=1000)*pi)
  y = sin(2*seq(0,pi,length.out=1000)*pi)
  plot(x,y,type="l",lwd=3,xlim=c(-1.1,1.1),ylim=c(-1.1,1.1),ann=F,axes=F)
  lines(x=c(-1,1,1,-1,-1)*1.1,y=c(1,1,-1,-1,1)*1.1,lwd=4)
  colors = c("yellow","yellow","red","red","blue","blue","grey20","grey20","white","white")
  colors2 = lapply(colors, col2rgb)
  linecolors = c(rep("black",6),"white",rep("black",3))
  textcolors = c(rep("black",6),rep("white",2),rep("black",2))
  for(i in 10:1){
    x2 = x*i/10
    y2 = y*i/10
    polygon(x2[1:(length(x2)-1)],y2[1:(length(x2)-1)],col=colors[i],border=F)
    lines(x2,y2,col=linecolors[i])
    text((-i+0.5)/10,0,11-i,col=textcolors[i],cex=1.5)
  }
  lines(x*0.5/10,y*0.5/10,col="grey50")
  points(0,0,pch=20,cex=0.5)
}
par(mar=c(0,0,0,0))
makebullseye()
```

# Modeling an archer

An archer can be modeled saying her points follow a uniform distribution.

```{r}
archer1 = sample(0:10,9999,replace=TRUE)
hist(archer1,breaks=0:11-0.5,freq=FALSE,xlab="points",main="An archer with pnts=uniform(0,10) , n=9999",col=rgb(0.8,0.3,0.3))
```
\

```{r,fig.width=6,fig.height=6}
par(mar=c(0,1,2,1))
makebullseye()
set.seed(123)
r = runif(999,0,1.1)
theta = runif(999,0,2*pi)
points(r*cos(theta),r*sin(theta),pch=4)#21,cex=0.3,lwd=1,bg=rgb(1,1,1),col="black")
mtext("Positions of n=999 shots from pnts=uniform(0,10) archer")
```
\

```{r}
archer1 = sample(6:10,9999,replace=TRUE)
hist(archer1,breaks=0:11-0.5,freq=FALSE,xlab="points",main="Point Histogram pnts=uniform(6,10), n=9999",col=rgb(0.8,0.3,0.3))
```
\

```{r,fig.width=6,fig.height=6}
par(mar=c(0,1,2,1))
makebullseye()
set.seed(123)
r = runif(99,0,0.5)
theta = runif(99,0,2*pi)
points(r*cos(theta),r*sin(theta),pch=4)#21,cex=0.3,lwd=1,bg=rgb(1,1,1),col="black")
mtext("Positions of n=999 shots from uniform(6,10) archer")
```

## position vs points

You may have noticed that the "10" score is harder to read than the other numbers in both positional graphs. 

A standard target has a diameter of 60 cm. This means each concentric circle has a radius that is 3 cm longer than its smaller circle.

```{r fig.width=6,fig.height=6}
makebullseye = function(scale=30,axes=T){
  x = cos(2*seq(0,pi,length.out=1000)*pi)*scale
  y = sin(2*seq(0,pi,length.out=1000)*pi)*scale
  plot(x,y,type="l",lwd=3,xlim=c(-scale,scale)*1.1,ylim=c(-scale,scale)*1.1,ann=F,axes=axes)
  lines(x=c(-1,1,1,-1,-1)*scale*1.1,y=c(1,1,-1,-1,1)*scale*1.1,lwd=4)
  colors = c("yellow","yellow","red","red","blue","blue","grey20","grey20","white","white")
  colors2 = lapply(colors, col2rgb)
  linecolors = c(rep("black",6),"white",rep("black",3))
  textcolors = c(rep("black",6),rep("white",2),rep("black",2))
  for(i in 10:1){
    x2 = x*i/10
    y2 = y*i/10
    polygon(x2[1:(length(x2)-1)],y2[1:(length(x2)-1)],col=colors[i],border=F)
    lines(x2,y2,col=linecolors[i])
  }
  lines(x*0.5/10,y*0.5/10,lty=3,col="grey50")
  points(0,0,pch=20,cex=0.5)
}
par(mar=c(3,3,3,3))
library(latex2exp)
makebullseye()
mtext("x, horizontal position (mm)",1,line=2)
mtext("y, vertical position (mm)",2,line=2)
mtext(TeX("$X=$unif(-30,30), $Y=$unif(-30,30), n=999"),3,line=1)
xarrows = runif(9999,-30,30)
yarrows = runif(9999,-30,30)
points(xarrows,yarrows,pch=4)#21,cex=0.3,lwd=1,bg=rgb(1,1,1),col="black")
```
\
```{r}
rarrows = sqrt(xarrows^2+yarrows^2)
ra2 = sapply(rarrows,function(x){min(x,30)})
pnts = 10-floor(ra2/3)
hist(pnts,breaks=0:11-0.5,freq=FALSE,xlab="points",main=TeX("$X=$unif(-30,30), $Y=$unif(-30,30), n=999"),col=rgb(0.8,0.3,0.3))
```
\

```{r fig.width=6,fig.height=6}
set.seed(123)
n=9999
par(mar=c(3,3,3,3))
makebullseye()
mtext("x, horizontal position (mm)",1,line=2)
mtext("y, vertical position (mm)",2,line=2)
mtext(TeX("$R=\\30\\sqrt{unif(0,1)}$; $\\theta = unif(0,2\\pi)$; $n=10$"),3,line=1)
mtext(TeX("$X= R\\cdot\\cos{\\theta}$; $Y= R\\cdot\\sin{\\theta}$"),3,line=0)
rarrowsB = sqrt(runif(n,0,1))*30
theta = runif(n,0,2*pi)
points(rarrowsB*cos(theta),rarrowsB*sin(theta),pch=4)
```
\

```{r}
hist(10-floor(rarrowsB/3),breaks=0:11-0.5,freq=FALSE,xlab="points",main=TeX("$R\\sim 30 \\sqrt{unif(0,1)}$; $\\theta\\sim unif(0,2\\pi)$"),col=rgb(0.8,0.3,0.3))
```
\


The surface areas of the regions are different. 
```{r}
tabstr = '|Points of region | Formula |  Area ($\\text{cm}^2$) |Proportion of area|\n|:---:|:---:|:---:|:---:|\n'
rs = 0:10*3
for(i in 1:10){
  tabstr = paste0(tabstr,"| ",11-i,"| $\\pi\\cdot",rs[i+1],"^2-\\pi\\cdot",rs[i],"^2$ | ",pi*(rs[i+1]^2-rs[i]^2),"|",pi*(rs[i+1]^2-rs[i]^2)/(pi*30^2),"|\n",collapse="")
}
```

`r tabstr`


$$(r+3)^2-r^2$$
$$r^2+6r+9-r^2$$
$$6r+9$$

## Chad's scores

```{r me}
pnts = sample(c(7,7,5,9,9,10,7,7,8,7,10,8))
chadpnts = c(10,10,9,8,8,8,7,6,5,4,10,9,8,8,8,7,6,6,6,4,9,9,8,8,7,7,5,5,4,1,9,9,9,8,8,8,7,6,5,5,9,8,8,7,7,6,6,5,5,4,8,8,8,7,7,7,6,6,5,3,9,9,9,8,8,7,7,6,6,2,10,10,9,9,7,7,5,4,4,1,10,9,8,8,6,6,6,6,6,4,9,9,8,7,7,6,6,5,2,2)
chadhist = hist(chadpnts,breaks=0:11-0.5,main="A sample of Chad's points, n=100",xlab="points",col=rgb(0.9,0.4,0.4),labels=T)
cnts = chadhist$counts
brks = chadhist$breaks
```
The average of Chad's points is $\bar{a}=`r mean(pnts)`$ and the standard deviation is $s=`r sd(pnts)`$.


```{r}
tabstr2 = '|Points of region | Frequency | Area (cm^2) | Positional Density (arrows/cm^2) |\n|:---:|:---:|:---:|:---:|\n'
rs = 0:10*3
for(i in 1:10){
  tabstr2 = paste0(tabstr2,"| ",11-i,"|",cnts[12-i],"|",pi*(rs[i+1]^2-rs[i]^2),"|",cnts[12-i]/(pi*(rs[i+1]^2-rs[i]^2)),"|\n",collapse="")
}
```
`r tabstr2`


```{r,fig.width=6,fig.height=6}
shots = read.csv("shots.csv")
set.seed(123)
chadx = shots$x/10*0.99
chady = shots$y/10*0.99
chadr = sqrt(chadx^2+chady^2)
chadt = atan(chady/chadx)*((chadx>0)-0.5)*2
par(mar=c(3,3,3,3))
makebullseye()
mtext("x, horizontal position (cm)",1,line=2)
mtext("y, vertical position (cm)",2,line=2)
mtext("Chad's 100 shots, positional",3,line=1)
points(chadx,chady,pch=4)
pntsC = 10-floor(chadr/3)
chadhist = hist(pntsC,breaks=0:10+0.5,col=rgb(0.9,0.3,0.3),labels=T)
cnts = chadhist$counts
```
```{r}
tabstr3 = '|Points of region | Frequency | Area (cm^2) | Positional Density (arrows/cm^2) |\n|:---:|:---:|:---:|:---:|\n'
rs = 0:10*3
for(i in 1:10){
  tabstr3 = paste0(tabstr3,"| ",11-i,"|",cnts[11-i],"|",pi*(rs[i+1]^2-rs[i]^2),"|",cnts[11-i]/(pi*(rs[i+1]^2-rs[i]^2)),"|\n",collapse="")
}
```
`r tabstr3`


```{r}
posden = numeric()
for(i in 1:10){
  posden = c(posden,cnts[11-i]/(pi*(rs[i+1]^2-rs[i]^2)))
}
barplot(posden)
x = seq(0,3,0.01)
plot(x,x^2*exp(-3*x))
```




```{r}
hist(chadx,breaks=seq(-33,33,6))
hist(chady,breaks=seq(-33,33,6))
plot(abs(chadx),abs(chady))
m1 = lm(abs(chady)~abs(chadx))
cor.test(chadx,chady)
cor.test(abs(chadx),abs(chady))
```

```{r}
t.test(chadx)
t.test(chady)
m2 = lm(chady~chadx)
m2
qqnorm(chadx)
```


Amelia shoots:
points = 9, 10, 9, 10.... xbar=9, sd=1






