---
header-includes:
   - \usepackage{arev}
   - \usepackage{multicol}
geometry: "left=1cm,right=1cm,top=1cm,bottom=1cm"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\pagestyle{empty} 
# Confidence Interval

* The confidence level, $\gamma$, represents how confident we are the interval will contain the population parameter (population proportion or population mean).
* To get $z^{\star}$, find $z^\star$ such that $P(|Z|<z^\star)=\gamma$. To do that, first get percentile, $\ell$, from confidence level ($\gamma$):
$$\ell = \frac{\gamma+1}{2} $$
then, use the $z$-table to find $z^\star$ such that $P(Z<z^\star)=\ell$.

### Proportion

The population proportion, $p$, is estimated with an interval (to indicate uncertainty) based on a sample proportion, $\hat{p}$.

* Bounds:
$$\hat{p}\pm z^{\star}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}} $$
* Necessary sample size for a given margin of error:  
    * If $\hat{p}$ is known:
      $$n = \hat{p}(1-\hat{p})\left(\frac{z^{\star}}{ME}\right)^2 $$
    * If $\hat{p}$ is unknown, assume it is 0.5 to be conservative
      $$n = \frac{1}{4}\left(\frac{z^{\star}}{ME}\right)^2 $$

### Mean

The population mean, $\mu$, is estimated with an interval (to indicate uncertainty) based on a sample mean, $\bar{x}$.

* Bounds:
    * If $\sigma$ is known:
      $$\bar{x} \pm z^{\star} \cdot \frac{\sigma}{\sqrt{n}}$$
    * If $\sigma$ is unknown, use the sample standard deviation (and $t^{\star}$). Remember, $df = n-1$. To get $t^{\star}$, find $t^\star$ such that $P(|T|<t^\star)=\gamma$ and $df=n-1$.
      $$\bar{x} \pm t^{\star} \cdot \frac{s}{\sqrt{n}} $$
* Necessary sample size for a given margin of error:
$$n = \left( \frac{z^{\star} \sigma}{ME} \right)^2 $$

\newpage

# Hypothesis Testing (Single-Sample) 

\vspace{-6pt}
\newcommand{\pval}{p\text{-value}}
\thispagestyle{empty}

\begin{center}
$\begin{aligned}
H_0 &= \text{null hypothesis} \\
H_A &= \text{alternative hypothesis} \\
\pval &= \text{probability of getting sample at least as extreme as observed sample, {\bf given} $H_0$} \\
\alpha &= \text{significance level = chance of type II error given $H_0$}
\end{aligned}$
\end{center}

\begin{itemize}
\item Calculate the $\pval$.
\begin{itemize}
\item "at least as extreme" can mean "as large or larger", "as small or smaller", or "as far from expected in either direction".
\end{itemize}
\item If $\pval$ is small enough, we reject the null hypothesis. (This logic is similar to {\it reductio ad absurdum} or  proof by contradiction.)
$$\text{If }~~\pval<\alpha ~~\text{ then }~~ \text{reject }H_0 $$
$$\text{If }~~\pval \ge \alpha ~~\text{ then }~~ \text{do not reject }H_0 $$
\end{itemize}

### Single-sample proportion testing

Necessary conditions: $\hat{p}n\ge 10$ and $(1-\hat{p})n\ge10$.

$$z_0 = \cfrac{\hat{p}-p_0}{\sqrt{\cfrac{p_0(1-p_0)}{n}}} $$

\begin{multicols}{3}
Right tail (one tail)
\begin{itemize}
\item $H_0$ claims $p = p_0$
\item $H_A$ claims $p > p_0$
\item $\pval = P(Z > z_0)$
\end{itemize}

Left tail (one tail)
\begin{itemize}
\item $H_0$ claims $p = p_0$
\item $H_A$ claims $p < p_0$
\item $\pval = P(Z < z_0)$
\end{itemize}

Two tail
\begin{itemize}
\item $H_0$ claims $p = p_0$
\item $H_A$ claims $p \ne p_0$
\item $\pval = P(|Z| > |z_0|)$
\end{itemize}
\end{multicols}

### Single-sample mean testing, $\sigma$ known

$$z_0 = \cfrac{\bar{x}-\mu_0}{\frac{\sigma}{\sqrt{n}}} $$

\begin{multicols}{3}
Right tail (one tail)
\begin{itemize}
\item $H_0$ claims $\mu = \mu_0$
\item $H_A$ claims $\mu > \mu_0$
\item $\pval = P(Z > z_0)$
\end{itemize}

Left tail (one tail)
\begin{itemize}
\item $H_0$ claims $\mu = \mu_0$
\item $H_A$ claims $\mu < \mu_0$
\item $\pval = P(Z < z_0)$
\end{itemize}

Two tail
\begin{itemize}
\item $H_0$ claims $\mu = \mu_0$
\item $H_A$ claims $\mu \ne \mu_0$
\item $\pval = P(|Z| > |z_0|)$
\end{itemize}
\end{multicols}

### Single-sample mean testing, $\sigma$ unknown

$$\begin{aligned}
df&=n-1 &&&&&&&&& t_0 &= \cfrac{\bar{x}-\mu_0}{\frac{s}{\sqrt{n}}} 
\end{aligned}$$

\begin{multicols}{3}
Right tail (one tail)
\begin{itemize}
\item $H_0$ claims $\mu = \mu_0$
\item $H_A$ claims $\mu > \mu_0$
\item $\pval = P(T > t_0)$
\end{itemize}

Left tail (one tail)
\begin{itemize}
\item $H_0$ claims $\mu = \mu_0$
\item $H_A$ claims $\mu < \mu_0$
\item $\pval = P(T < t_0)$
\end{itemize}

Two tail
\begin{itemize}
\item $H_0$ claims $\mu = \mu_0$
\item $H_A$ claims $\mu \ne \mu_0$
\item $\pval = P(|T| > |t_0|)$
\end{itemize}
\end{multicols}









\newpage
```{r getlatex}
library(latex2exp)
`%+%` <- function(x,y)
{
return(paste(x,y,sep=""))
}
tab = ''
tab = tab %+% paste(rep("$z$ & $P(Z<z)$",6), collapse=" & ")
tab = tab %+% '\\\\ \\hline\n'

for(i in seq(0,0.5,0.01))
{
    zs = seq(-3,-0.5,0.5)+i
    Ps = round(pnorm(zs),4)
    zs = sprintf("%0.2f", zs)
    Ps = sprintf("%0.4f", Ps)
    line = paste(zs,Ps,sep = " & ", collapse= " & ")
    tab = tab %+%  paste(line,'\\\\\n',collapse = "" )
}



tab2 = paste(rep("$z$ & $P(Z<z)$",6), collapse=" & ")
tab2 = tab2 %+% '\\\\ \\hline\n'

for(i in seq(0,0.5,0.01))
{
    zs = seq(0,2.5,0.5)+i
    Ps = round(pnorm(zs),4)
    zs = " " %+% sprintf("%0.2f", zs)
    Ps = sprintf("%0.4f", Ps)
    line = paste(zs,Ps,sep = " & ", collapse= " & ")
    tab2 = tab2 %+% paste(line,'\\\\\n',collapse = "" )
}
```

\thispagestyle{empty}

\vspace{-20pt}
{\LARGE z-table}$~$\hfill
```{r negnorm,fig.height=1.2,fig.width=4}
z = (-300:300)/100
par(mar=c(2,0,0,0))
plot(z,dnorm(z),type="l",axes=F,ann=F,xlim=c(-3,3),ylim=c(0,0.4))
axis(1,at=-3:3)
polygon(c(z[z<=-1.4],-1.4,-3),c(dnorm(z[z<=-1.4]),0,0),col=rgb(1,0,0,0.4))
mtext(TeX("$z$"),1,line=0,at=-1.4)
text(-2.0,0.35,TeX("left area  =  $P(Z$ $<$ $z)$"))
for(x in seq(-3,3,0.2)){
    lines(c(x,x),c(0,dnorm(x)),lty=3,lwd=0.25)
}
for(y in seq(0,0.35,0.05)){
    xmin = min(z[dnorm(z)>=y])
    xmax = max(z[dnorm(z)>=y])
    lines(c(xmin,xmax),c(y,y),lty=3,lwd=0.25)
}
```


\begin{center}
\begin{tabular}{|c c|c c|c c|c c|c c|c c|}
`r tab`
\end{tabular}
\end{center}

\newpage

\vspace{-20pt}
{\LARGE z-table}$~$\hfill
```{r posnorm,fig.height=1.2,fig.width=4}
library(latex2exp)
z = (-300:300)/100
par(mar=c(2,0,0,0))
plot(z,dnorm(z),type="l",axes=F,ann=F,xlim=c(-3,3),ylim=c(0,0.4))
axis(1,at=-3:3)
polygon(c(z[z<=1.4],1.4,-3),c(dnorm(z[z<=1.4]),0,0),col=rgb(1,0,0,0.4))
mtext(TeX("$z$"),1,line=0,at=1.4)
text(-2,0.35,TeX("left area  =  $P(Z$ $<$ $z)$"))
for(x in seq(-3,3,0.2)){
    lines(c(x,x),c(0,dnorm(x)),lty=3,lwd=0.25)
}
for(y in seq(0,0.35,0.05)){
    xmin = min(z[dnorm(z)>=y])
    xmax = max(z[dnorm(z)>=y])
    lines(c(xmin,xmax),c(y,y),lty=3,lwd=0.25)
}
```

\thispagestyle{empty}

\begin{center}
\begin{tabular}{|c c|c c|c c|c c|c c|c c|}
`r tab2`
\end{tabular}
\end{center}

\newpage


```{r getlatext}
library(latex2exp)

`%+%` <- function(x,y)
{
    return(paste(x,y,sep=""))
}

las = c(0.90,0.95,0.975,0.98,0.99,0.995,0.9975,0.998,0.999)
ras = c(0.1,0.05,0.025,0.02,0.01,0.005,0.0025,0.002,0.001)
mas = c(0.8,0.9,0.95,0.96,0.98,0.99,0.995,0.996,0.998)
tts = c(0.2,0.1,0.05,0.04,0.02,0.01,0.005,0.004,0.002)

tab = "&& $P(T<t)=$ & "
tab = tab %+% paste(las, collapse=" & ")
tab = tab %+% "\\\\\n"
tab = tab %+% "&& $P(T>t)=$ & "
tab = tab %+% paste(ras, collapse=" & ")
tab = tab %+% "\\\\\n"
tab = tab %+% "&& $P(|T|<t)=$ & "
tab = tab %+% paste(mas, collapse=" & ")
tab = tab %+% "\\\\\n"
tab = tab %+% "&& $P(|T|>t)=$ & "
tab = tab %+% paste(tts, collapse=" & ")
tab = tab %+% "\\\\\\hline\n"

tab = tab %+% "df= && $t=$ & &&&&&&&& "
tab = tab %+% "\\\\ \n"

tab2 = tab

dfs = 1:40

for (df in dfs)
{
    tab = tab %+% "&"
    tab = tab %+% df
    tab = tab %+% "&& "
    tab = tab %+% paste(sprintf("%.2f",qt(las,df),2), collapse=" & ")
    tab = tab %+% "\\\\ \n"
}

dfs = c(41:50,seq(55,100,5),seq(125,500,25))

for (df in dfs)
{
    tab2 = tab2 %+% "&"
    tab2 = tab2 %+% df
    tab2 = tab2 %+% "&& "
    tab2 = tab2 %+% paste(sprintf("%.2f",qt(las,df),2), collapse=" & ")
    tab2 = tab2 %+% "\\\\ \n"
}

tab2 = tab2 %+% "&"
tab2 = tab2 %+% ("$\\infty$")
tab2 = tab2 %+% "&& "
tab2 = tab2 %+% paste(sprintf("%.2f",qnorm(las),2), collapse=" & ")
tab2 = tab2 %+% "\\\\ \n"

```
\pagestyle{empty}
\large

T-table
\renewcommand{\arraystretch}{1.1}
\begin{center}
\begin{tabular}{|c c| l c c c c c c c c c|}\hline
`r tab` \hline
\end{tabular}
\end{center}

\newpage

T-table
\renewcommand{\arraystretch}{1.1}
\begin{center}
\begin{tabular}{|c c| l c c c c c c c c c|}\hline
`r tab2` \hline
\end{tabular}
\end{center}