---
title: "stdevp"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Simple spinner (equally sized wedges)

* A simple spinner is split into equally sized wedges. 
* As the spinner is the source of the data, it is the "population".
* The symbol for population mean is $\mu$ ("mu").
* The symbol for population standard deviation is $\sigma$ ("sigma").
* In these slides, we will use $N$ (upper case) as the number of equally-sized wedges.
* In these slides, we will use $X$ (upper case) as the list of values on the wedges.

```{r simpspin,fig.width=1.5,fig.height=1.5}
draw_simp_spinner = function(X){
  N = length(X)
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",axes=F,ann=F)
  polygon(cos(seq(0,2*pi,length.out=1000)), sin(seq(0,2*pi,length.out=1000)))
  for(i in 1:N){
    lines(c(0,cos((i-1)/N*2*pi)),c(0,sin((i-1)/N*2*pi)))
    text(0.75*cos((i-0.5)/N*2*pi),0.75*sin((i-0.5)/N*2*pi),X[i])
  }
  arrows(0.5*cos(pi+0.1),0.5*sin(pi+0.1),0.5*cos(0.1),0.5*sin(0.1),length=0.1,code=2,lwd=2)
}

X = c(2,5,8,20,2,11)
N = length(X)
draw_simp_spinner(X)
```


## Population mean

```{r simpspin2,fig.width=1.5,fig.height=1.5}
draw_simp_spinner(X)
```

From a simple spinner, the population mean can be found by summing the values and dividing by the number of wedges.
$$\mu = \frac{\sum X}{N} = \frac{2+5+8+20+2+11}{6} = `r mean(X)` $$

## Population standard deviation

The population standard deviation uses a similar formula as the sample standard deviation, but there is not Bessel correction.
$$\sigma = \sqrt{\frac{\sum (X-\mu)^2}{N}} $$
You can just use geogebra...
$$\texttt{stdevp(2,5,8,20,2,11)} = `r sd(X)*sqrt((N-1)/N)` $$
(notice the "p" at the end of "stdevp" stands for population)


## total of 2 spins

Consider the following spinner ($X$):

```{r three,fig.width=1,fig.height=1}
X = c(1,2,3)
N=length(X)
draw_simp_spinner(X)
```

Now, consider the possibilities of spinning it twice and adding the results. Each of the following sequences would be equally likely:
(1,1), (1,2), (1,3), (2,1), (2,2), (2,3), (3,1), (3,2), (3,3)

Thus, we could imagine the following spinner representing $X+X$:

```{r threetwice,fig.width=1,fig.height=1}
X = c(2,3,4,3,4,5,4,5,6)
N=length(X)
draw_simp_spinner(X)
```

## total of 2 spins

```{r three2,fig.width=1,fig.height=1,fig.align='center'}
X1 = c(1,2,3)
N1=length(X1)
draw_simp_spinner(X1)
mu1 = mean(X1)
sig1 = sd(X1)*sqrt((N1-1)/N1)
```
\vspace{-30pt}
$$\mu = `r mean(X1)` $$
$$\sigma = `r sd(X1)*sqrt((N1-1)/N1)` $$

```{r threetwice2,fig.width=1,fig.height=1,fig.align='center'}
X2 = c(2,3,4,3,4,5,4,5,6)
N2=length(X2)
draw_simp_spinner(X2)
mu2 = mean(X2)
sig2 = sd(X2)*sqrt((N2-1)/N2)
```
\vspace{-30pt}
$$\mu = `r mu2` $$
$$\sigma = `r sig2` $$


## total of 2 spins

Notice the mean doubled, but standard deviation did not.
$$\frac{4}{2} = 2 $$

$$\frac{`r sig2`}{`r sig1`} = `r sig2/sig1` = \sqrt{2} $$

## Doubling 1 spin

Consider the following spinner ($X$):

```{r threeA,fig.width=1,fig.height=1}
X = c(1,2,3)
N=length(X)
draw_simp_spinner(X)
```

Now, consider the possibilities of spinning it once and doubling the results. Each of the following outcomes would be equally likely:
2, 4, 6

Thus, we could imagine the following spinner representing $2X$:

```{r threetwiceA,fig.width=1,fig.height=1}
X = c(2,4,6)
N=length(X)
draw_simp_spinner(X)
```

## Doubling 1 spin

```{r three2A,fig.width=1,fig.height=1,fig.align='center'}
X1 = c(1,2,3)
N1=length(X1)
draw_simp_spinner(X1)
mu1 = mean(X1)
sig1 = sd(X1)*sqrt((N1-1)/N1)
```
\vspace{-30pt}
$$\mu = `r mean(X1)` $$
$$\sigma = `r sd(X1)*sqrt((N1-1)/N1)` $$

```{r threetwice2A,fig.width=1,fig.height=1,fig.align='center'}
X2 = c(2,4,6)
N2=length(X2)
draw_simp_spinner(X2)
mu2 = mean(X2)
sig2 = sd(X2)*sqrt((N2-1)/N2)
```
\vspace{-30pt}
$$\mu = `r mu2` $$
$$\sigma = `r sig2` $$

## Doubling 1 spin

Notice that both the mean and standard deviation doubled.

$$\frac{`r sig2`}{`r sig1`} = `r sig2/sig1` $$

## Theory (linear combination of random variables)

* If $X$ and $Y$ represent two random variables, and $a$ and $b$ represent two constants, then:
$$SD(aX+bY) = \sqrt{a^2 SD(X)^2+b^2 SD(Y)^2} $$
$$SD(X+Y) = \sqrt{SD(X)^2+SD(Y)^2} $$
$$SD(X+X) = \sqrt{SD(X)^2+SD(X)^2} = \sqrt{2SD(X)^2}= \sqrt{2}SD(X)$$
$$SD(X+X+X) = \sqrt{SD(X)^2+SD(X)^2+SD(X)^2}$$
$$SD(aX) = \sqrt{a^2 SD(X)^2} = a \cdot SD(X)$$
