```{r data generation, echo = FALSE, results = "hide"}
while(T){
  ell = sort(runif(2,0.02,0.98))
  z = round(qnorm(ell),2)
  ell = round(pnorm(z),4)
  sol = ell[2]-ell[1]
  if(z[2]-z[1] > 0.5){break}
}
```

Question
========
Evaluate $P(`r z[1]` < Z < `r z[2]`)$.


Solution
========
Start with a sketch.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=z[1] & zs<=z[2]],z[2],z[1]),
        c(dnorm(zs[zs>=z[1] & zs<=z[2]]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text((z[1]+z[2])/2,dnorm((z[1]+z[2])/2)/3,"?")
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z[1],z[2],3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\

We take a difference of areas.
$$\begin{aligned}
P(`r z[1]`<Z<`r z[2]`) &= P(Z<`r z[2]`) - P(Z<`r z[1]`) \\
&= `r ell[2]` - `r ell[1]` \\
&= `r sol`
\end{aligned}$$

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=z[1] & zs<=z[2]],z[2],z[1]),
        c(dnorm(zs[zs>=z[1] & zs<=z[2]]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text((z[1]+z[2])/2,max(dnorm((z[1]+z[2])/2)/3,0.02),sol)
text((z[1]-3)/2,max(dnorm((z[1]-3)/2)/3,0.02),ell[1])
text((-3+z[2])/2,0.46,ell[2])
lines(c(-3,3),c(0,0))
lines(c(-3,z[2],z[2]),c(0.43,0.43,0.41))
axis(1,at=c(-3,z[1],z[2],3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\

Meta-information
================
extype: num
exsolution: `r sol`
exname: Get area between 2 zs
extol: 0.0002
