```{r data generation, echo = FALSE, results = "hide"}
ell = runif(1,0.65,0.98)
z = round(qnorm(ell),2)
ell = round(pnorm(z),4)
sol = (ell-0.5)*2

la = round(pnorm(-z),4)
```

Question
========
Evaluate $P\left(|Z| < `r z`\right)$.


Solution
========
First, you need to identify that we are looking for a central area. This is because $z$-scores near 0 satisfy $|Z|<`r z`$, while $z$-scores far from 0 (either positive or negative) do not satisfy the inequality.

Start with a sketch.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=-z & zs<=z],z,-z),c(dnorm(zs[zs>=-z & zs<=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
text(0,0.2,"?")
```
\

Remember, the entire area is always 1. Our z-table only provides left areas. 

**Method 1:** First find the area of the left tail.
$$P(Z<`r -z`) =  `r la`$$
Recognize normal distributions are symmetric, so we also know the area of the right tail.
$$P(Z>`r z`) = `r la` $$
The three areas add to 1.
$$\begin{aligned}
P\left(|Z|<`r z`\right) &= 1-P(Z<`r -z`)-P(Z>`r z`) \\
&= 1 - `r la` - `r la` \\
&= `r sol`
\end{aligned}$$

This technique can be summarized with the following formula:
$$P\left(|Z|<z\right) = 1-2\cdot P(Z<-z)$$
assuming $z>0$.

So, in our case (when $z=`r z`$),
$$\begin{aligned}
P\left(|Z|<`r z`\right) &= 1-2\cdot P(Z<`r -z`) \\
&= 1-2\cdot`r la` \\
&= `r sol`
\end{aligned}$$

This method is shown graphically:

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=-z & zs<=z],z,-z),c(dnorm(zs[zs>=-z & zs<=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
text(0,0.2,sol)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/2,0.01),la)
text(xr,max(dnorm(xr)/2,0.01),la)
```
\

**Method 2:** You can also find half of the central area and then double it.
$$\begin{aligned}
P\left(|Z|<z\right) &= 2\cdot\left[ P(Z<z)-0.5\right] \\
&= 2\cdot\left[ `r ell`-0.5\right] \\
&= 2\cdot\left[ `r ell-0.5`\right] \\
&= `r sol`
\end{aligned}$$

This method is shown graphically:

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=-z & zs<=z],z,-z),c(dnorm(zs[zs>=-z & zs<=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(0,0),c(0,dnorm(0)))
lines(c(-3,z,z),c(0.42,0.42,0.4))
text((-3+z)/2,0.45,ell)
text(z/2,0.1,ell-0.5,cex=0.6)
```
\

**Method 3:** You can also calculate the central area with a difference of two left areas.

$$\begin{aligned}
P\left(|Z|<`r z`\right) &= P(Z<`r z`)-P(Z<`r -z`)\\
&= `r ell`-`r la` \\
&= `r sol`
\end{aligned}$$

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=-z & zs<=z],z,-z),c(dnorm(zs[zs>=-z & zs<=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(-3,z,z),c(0.42,0.42,0.4))
text((-3+z)/2,0.45,ell)
text(0,0.1,sol)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/2,0.02),la)
```
\

Meta-information
================
extype: num
exsolution: `r sol`
exname: Get central area from Z
extol: 0.0002
