```{r data generation, echo = FALSE, results = "hide"}
library(latex2exp)
while(T)
{
  ell = round(runif(1,0.60,0.98),2)
  ra = 1-ell
  ca = 1-2*ra
  ta = ra*2
  z = qnorm(ell)
  zlow = floor(z*100)/100
  zhigh = ceiling(z*100)/100
  if(round(z,4) != round(z,2)) {break}
}
```

Question
========
Determine $z$ such that $P(|Z|>z)=`r ta`$.

Solution
========
Start with a sketch. The total two-tail area is `r ta`, so each tail has half that area.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),
        c(dnorm(zs[zs<=-z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text(min(z+0.4,2.6),max(dnorm(z+0.4)/3,0.04),TeX(sprintf("$\\frac{%.2f}{2}$",ta)))
text(max(-z-0.4,-2.6),max(dnorm(z+0.4)/3,0.04),TeX(sprintf("$\\frac{%.2f}{2}$",ta)))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","z=?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\

**Method 1:** Determine the area of each tail and the center. Both tails have the same area, and all three areas add to 1. Thus,
$$P(Z>z) = \frac{`r ta`}{2} = `r ra`$$

$$P(Z<-z) = \frac{`r ta`}{2} = `r ra`$$

$$\begin{aligned}
P(|Z|<z) &= 1-P(|Z|>z) \\ 
&= 1-`r ta` \\
&= `r ca`
\end{aligned}$$
```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),
        c(dnorm(zs[zs<=-z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text(0,dnorm(0)/3,ca)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/3,0.02),ra)
text(xr,max(dnorm(xr)/3,0.02),ra)
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","z=?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(-3,z,z),c(0.42,0.42,0.4))
text((-3+z)/2,0.45,ell)
```
\

Find the left area.

$$\begin{aligned}
P(Z<z) &= `r ra`+`r ca` \\
&= `r ell`
\end{aligned}$$


```{r, echo=F}
z2 = seq(zlow-0.02,zhigh+0.02,0.01)
```

You should go to your $z$-table and find the $z$-score with the left area closest to `r ell`.

| $z$ | $P(Z<z)$ |
|:---:|:---:|
| `r z2[1]` | `r format(round(pnorm(z2[1]),4),4)` |
| `r z2[2]` | `r format(round(pnorm(z2[2]),4),4)` |
| **`r z2[3]`** | **`r format(round(pnorm(z2[3]),4),4)`** |
| **`r z2[4]`** | **`r format(round(pnorm(z2[4]),4),4)`** |
| `r z2[5]` | `r format(round(pnorm(z2[5]),4),4)` |
| `r z2[6]` | `r format(round(pnorm(z2[6]),4),4)` |

It turns out the exact answer is `r z`, which could be found by using an inverse normal function. Like InvNorm(`r ell`) on a TI calculator, Norm.Inv(`r ell`) on a spreadsheet, or qnorm(`r ell`) in R.

But, because we are using the $z$-table, I will accept either `r zlow` or `r zhigh`. (Or, really anything within 0.011 of `r z`.)

**Method 2:** Another way to get `r ell` is by subtracting half of `r ta` from 1.
$$P(Z<z) ~=~ 1-\frac{`r ta`}{2}~=~ 1-`r ra` ~=~ `r ell`$$

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),
        c(dnorm(zs[zs<=-z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text(min(z+0.4,2.6),max(dnorm(z+0.4)/3,0.04),ra)
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","z=?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(-3,z,z),c(0.46,0.46,0.45))
text((-3+z)/2,0.476,ell)
```
\

Then, use the table.

Meta-information
================
extype: num
exsolution: `r z`
exname: Get z from twotail area
extol: 0.011
