```{r data generation, echo = FALSE, results = "hide"}
mux = sample(460:1200/10,1)
n = sample((6:17)^2,1)
sigx = signif(runif(1,mux/10/sqrt(n),mux/3.5/sqrt(n)),2)
muw = mux
sigw = sigx*sqrt(n)

while(T){
  p = sort(round(runif(2,0.02,0.98),3))
  if(p[2]-p[1] > 0.2){break}
}
z = round(qnorm(p),2)
x = signif(mux+z*sigx,3)
z = round((x-mux)/sigx,2)
p = round(pnorm(z),4)

P1 = p[1]
P2 = 1-p[2]
P3 = p[2]-p[1]
P = c(P1,P2,P3)
```

Question
========

A very large population has a mean of `r muw` and a standard deviation of `r sigw`. When a random sample of size `r n` is taken, what is the probability that the **sample mean** ($\bar{x}$) is between `r x[1]` and `r x[2]`?


Answerlist
--------

* Determine the mean and standard deviation of the sampling distribution. \newline\vfill
* Determine $P(`r x[1]` < \overline{X} < `r x[2]`)$. **Draw a sketch** \newline\vfill


Solution
========

\setkeys{Gin}{width=.5\textwidth}

Answerlist
--------
* Central limit of average formulas: $\mu_{\tiny \bar{x}} = `r mux`$ and $\sigma_{\tiny \bar{x}} = \frac{`r sigw`}{\sqrt{`r n`}} = `r sigx`$.
* $P(`r x[1]` < \overline{X} < `r x[2]`) ~=~ P(`r z[1]` < Z < `r z[2]`) ~=~ `r P3`$
\newline
```{r no7,echo=F,fig.height=2,fig.width=7,fig.cap=""}
library(latex2exp)
par(mar=c(4,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
Z = -350:350/100
lines(Z,dnorm(Z))
abline(h = 0)
axis(1,-3:3,labels=(-3:3)*sigx+mux,line=2, tick = FALSE)
mtext(TeX("$\\bar{x}=$"),1,3,adj = 0)
polygon(c(Z[Z>=z[1] & Z<=z[2]],z[2],z[1]),c(dnorm(Z[Z>=z[1] & Z<=z[2]]),0,0),col=rgb(0.4,0.4,0.4,0.4))
```
\setkeys{Gin}{width=\textwidth}


Meta-information
============
extype: string
exsolution: yo
exname: marbles
extol: 0.01
