```{r data generation, echo = FALSE, results = "hide", fig.show="hide"}
## DATA GENERATION
while(T){
  n = sample(4:7,1)
  mu = sample(40:200,1)
  sigma = sample(3:12,1)
  x = round(rnorm(n,mu,sigma))
  s = sd(x)
  if(as.integer(s)==s & s!=0){
    break    
  }
}
xbar = mean(x)
deviations = x-xbar
squared_deviations = deviations^2
sum_squared_deviations = sum(squared_deviations)
variance = sum_squared_deviations/(n-1)
ss = sqrt(variance)
```

Question
========

From a very large population, a small sample of measurements was taken.

`r knitr::kable(matrix(x,ncol=n))`

Please calculate the (Bessel corrected) sample standard deviation using the following formula:

$$s = \sqrt{\frac{\sum(x-\bar{x})^2}{n-1}}$$



Solution
========

```{r readyTable, echo=F}
tab1 = ''
for(i in 1:n){
  tab1 = paste(tab1,"|",x[i],"|",deviations[i],"|",squared_deviations[i],"|",sep="")
  if(i<n){
    tab1 = paste(tab1,"\n",sep="")
  }
}
```

We fill out the table column by column.

| $x$ | $x-\bar{x}$ | $(x-\bar{x})^2$ |
|:---:|:---:|:---:|
`r tab1`
|========|========|========|
|$\sum x = `r sum(x)`$| | $\sum (x-\bar{x})^2 = `r sum_squared_deviations`$|
|$\bar{x} = `r xbar`$| | |

We are ready for the formula.
$$\begin{aligned}
s &= \sqrt{\frac{\sum(x-\bar{x})^2}{n-1}} \\\\
&= \sqrt{\frac{`r sum_squared_deviations`}{`r n`-1}} \\\\
&= \sqrt{`r sum_squared_deviations/(n-1)`} \\\\
&= \fbox{`r sd(x)`}
\end{aligned}$$



Meta-information
============
extype: num
exsolution: `r sd(x)`
exname: calc sd