---
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\thispagestyle{fancy}

<!-- ## Instructions -->

<!-- This test is to be taken as an individual without outside assistance or notes. If you are suspected of cheating, you will fail this exam and your transgression will be reported. -->

<!-- You can either use a scientific calculator or (with a smartphone) GeoGebra Scientific Calculator. -->

<!-- To use GeoGebra, you must first put your smartphone on **Airplane Mode**. Then, in GeoGebra, turn on **Exam Mode**. You must leave exam mode on for the entire exam. You won't be able to use your smartphone for anything else. After you are done, you will show me how long exam mode has been running, and if that time is not how long you've been sitting, you will fail this exam. -->

<!-- Read each question carefully and show your work. -->

<!-- ## Grade Table -->

<!-- (do not write in this table) -->

<!-- \Large  -->

<!-- \begin{tabular}{|c|c|c|} \hline -->
<!-- question & available points & earned points \\ \hline \hline -->
<!-- 1 & 20 & \\ \hline -->
<!-- 2 & 20 &  \\ \hline -->
<!-- 3 & 10 &  \\ \hline -->
<!-- 4 & 10 &  \\ \hline -->
<!-- 5 & 10 &  \\ \hline -->
<!-- 6 & 10 &  \\ \hline -->
<!-- 7 & 20 & \\ \hline -->
<!-- EC1 & 10 &  \\ \hline -->
<!-- Total & 100 & \\ \hline -->
<!-- \end{tabular} -->

<!-- \normalsize -->

<!-- \newpage -->
\chead{}
\rhead{Formulas}

### Normal standardization
$$\begin{aligned}
z &= \frac{x-\mu}{\sigma} &&&&& x &= \mu + z\sigma
\end{aligned}$$

### Central Limit Theorem

- **If**:
  - $W$ is a random variable with mean $\mu_{\tiny w}$ and SD $\sigma_{\tiny w}$.
  - Random variable $X$ is **sum** of $n$ instances of $W$. $$X = W_1+W_2+W_3+\cdots+W_n $$
  - Random variable $Y$ is **average** of $n$ instances of $W$. $$Y = \frac{W_1+W_2+W_3+\cdots+W_n}{n} $$
- **Then**:
$$\begin{aligned} 
\mu_{\tiny x} &= n \mu_{\tiny w} &&&&& \sigma_{\tiny x} &= \sigma_{\tiny w} \sqrt{n} \\
\mu_{\tiny y} &= \mu_{\tiny w} &&&&& \sigma_{\tiny y} &= \frac{\sigma_{\tiny w}}{\sqrt{n}} 
\end{aligned}$$
  - $X$ and $Y$ are approximately normal (even if $W$ is not normal)

### Special case of central limit theorem: Bernoulli, Binomial, and $\hat{p}$ sampling

- **Given:**
  - $W$ is a Bernoulli random variable:
    \newline\begin{center}\begin{tabular}{|c|c|}\hline
    $w$ & $P(w)$ \\ \hline
    0 & $q$ \\
    1 & $p$ \\ \hline
    \end{tabular}\end{center}
  - $X$ is a binomial random variable (sum of $n$ instances of $W$)
  - $\hat{p}$ is a sample proportion (average of $n$ instances of $W$)
- **Therefore:**
$$\begin{aligned} 
\mu_{\tiny w}&=p &&&&& \sigma_{\tiny w}&=\sqrt{pq} \\
\mu_{\tiny x} &= np &&&&& \sigma_{\tiny x} &= \sqrt{pq}\sqrt{n} \\
\mu_{\tiny \hat{p}} &= p &&&&& \sigma_{\tiny \hat{p}} &= \frac{\sqrt{pq}}{\sqrt{n}} 
\end{aligned}$$
  

\newpage


## Question 1

\rhead{}

```{r data generation, echo = FALSE, results = "hide"}
set.seed(13)
types = c("A","B","C","D")
while(T){
  mus = sample(60:150,4)
  sigs = sample(4:15,4)
  ps = round(runif(4,0.02,0.98),4)
  zs = round(qnorm(ps),2)
  xs = signif(mus+zs*sigs,4)
  zs = round((xs-mus)/sigs,2)
  ps = round(pnorm(zs),4)
  if(sum(sort(zs)==sort(unique(zs)))==4 & sum(zs<0)>1){
    break
  }
}
```

A farm produces 4 types of fruit: $A$, $B$, $C$, and $D$. The fruits' masses follow normal distributions, with parameters dependent on the type of fruit.

\begin{center}
\begin{tabular}{|c|c|c|}\hline
Type of fruit & Mean mass (g) & Standard deviation of mass (g) \\ \hline
$A$ & `r mus[1]` & `r sigs[1]` \\
$B$ & `r mus[2]` & `r sigs[2]` \\
$C$ & `r mus[3]` & `r sigs[3]` \\
$D$ & `r mus[4]` & `r sigs[4]` \\ \hline
\end{tabular}
\end{center}

One specimen of each type is weighed. The results are shown below.

\begin{center}
\begin{tabular}{|c|c|}\hline
Type of fruit & Mass of specimen (g) \\ \hline
$A$ & `r xs[1]` \\
$B$ & `r xs[2]` \\
$C$ & `r xs[3]` \\
$D$ & `r xs[4]` \\ \hline
\end{tabular}
\end{center}

Which specimen is the most unusually small (relative to others of its type)?


\newpage

## Question 1 solution

We compare the $z$-scores. The smallest $z$-score corresponds to the specimen that is most unusually small.

\begin{center}
\begin{tabular}{|c|c|c|}\hline
Type of fruit & formula &  $z$-score \\ \hline
$A$ & $z=\frac{`r xs[1]`-`r mus[1]`}{`r sigs[1]`}$ &`r zs[1]` \\
$B$ & $z=\frac{`r xs[2]`-`r mus[2]`}{`r sigs[2]`}$ &`r zs[2]` \\
$C$ & $z=\frac{`r xs[3]`-`r mus[3]`}{`r sigs[3]`}$ &`r zs[3]` \\
$D$ & $z=\frac{`r xs[4]`-`r mus[4]`}{`r sigs[4]`}$ &`r zs[4]` \\ \hline
\end{tabular}
\end{center}

Thus, the specimen of type $`r types[zs==min(zs)]`$ is the most unusually small.


\newpage

## Question 2

```{r data generation 2, echo = FALSE, results = "hide"}
set.seed(2)
mu = sample(46:120,1)
sig = sample(3:(mu%/%3),1)
p1 = round(runif(1,0.02,0.98),3)
p2 = round(runif(1,0.02,0.98),3)
p3 = round(runif(1,0.52,0.98),3)
p4 = round(runif(1,0.52,0.98),3)
p = c(p1,p2,p3,p4)

z = round(qnorm(p),1)
p = round(pnorm(z),4)
P1 = p[1]
P2 = 1-p[2]
P3 = p[3]*2-1
P4 = (1-p[4])*2
P = c(P1,P2,P3,P4)
x = signif(mu+z*sig,3)
a = x-mu
```

Let $X$ be normally distributed with mean `r mu` and standard deviation `r sig`. Please calculate the probabilities shown below and also shade a corresponding region under the density curve.

\setkeys{Gin}{width=.6\textwidth}

a. $P(X<`r x[1]`)$  \newline \vfill \hfill
```{r no1,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
```
\vfill
b. $P(X>`r x[2]`)$  \newline \vfill \hfill
```{r no2,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
lines(-35:35/10,dnorm(-35:35/10))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
``` 
\vfill


\newpage

## Question 2 solution

a. $P(X<`r x[1]`) ~=~ P(Z<`r z[1]`) ~=~ \fbox{`r P1`}$ \newline \vfill
```{r no5,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
Z = -350:350/100
lines(Z,dnorm(Z))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
polygon(c(Z[Z<=z[1]],z[1],-3.5),c(dnorm(Z[Z<=z[1]]),0,0),col=rgb(0.4,0.4,0.4,0.4))
```
\vfill
b. $P(X>`r x[2]`) ~=~ P(Z>`r z[2]`) ~=~ \fbox{`r P2`}$ \newline \vfill
```{r no6,echo=F,fig.height=2,fig.width=7,fig.cap=""}
par(mar=c(2,0,0,0))
plot(0,0,type="n",xlim=c(-3.5,3.5),ylim=c(0,0.42),ann=F,axes=F)
axis(1,-3:3)
mtext("z=",1,1,adj = 0)
Z = -350:350/100
lines(Z,dnorm(Z))
for(xx in seq(-3.4,3.4,by=0.2)){
  abline(v = xx,lty=2)
}
for(yy in seq(0,0.4,by=0.05)){
  abline(h=yy,lty=2)
}
polygon(c(Z[Z>=z[2]],3.5,z[2]),c(dnorm(Z[Z>=z[2]]),0,0),col=rgb(0.4,0.4,0.4,0.4))
```
\vfill

\newpage



