---
title: "Spinner homework"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
set.seed(12345)
```


## Homework due Feb 11th

* Pick a spinner or other random number generator ($X$)
    - Show or describe the spinner in the report
* Plot a running average with at least 10 spins
    - Also, make a table showing the spins and calculations
    - Also, **describe** what you think the running average would do if you continued with more spins.
* Sample from $X+X+X$ at least 10 times; determine sample mean and sample standard deviation.
    - Also, make a table showing the measurements and calculations
* Sample from $3X$ at least 10 times; determine the sample mean and sample standard deviation.
    - Also, make a table showing the measurements and calculations
* In your own words, explain why $X+X+X$ has a smaller standard deviation than $3X$.
* Combine all work, with descriptions/explanations, into a report. Staple the pages together.

## Pick a spinner (or any random number generator)

* Pick a spinner or some other random number generator
* Some options (pick one):
    - Make your own spinner
    - Use a die (or dice) (directly or with a transformation table)
    - Use a provided spinner
    - Use a computer's random number generator
        - Spreadsheets have $\texttt{rand()}$ function, which generates standard uniform measurements
        - R has various functions: $\texttt{sample()}$, $\texttt{rnorm()}$, $\texttt{runif()}$, $\texttt{rgeom()}$, ...
* Examples:
```{r example,fig.width=1,fig.height=1}
X = c(0,0,0,1,1,1,1,1)
N = length(X)
par(mar=c(0,0,0,0))
plot(0,0,type="n",axes=F,ann=F)
polygon(cos(seq(0,2*pi,length.out=1000)), sin(seq(0,2*pi,length.out=1000)))
for(i in 1:N){
  lines(c(0,cos((i-1)/N*2*pi)),c(0,sin((i-1)/N*2*pi)))
  text(0.75*cos((i-0.5)/N*2*pi),0.75*sin((i-0.5)/N*2*pi),X[i])
}
```
```{r example2,fig.width=1,fig.height=1}
X = c(10,12,14,15,16,18,20)
N = length(X)
par(mar=c(0,0,0,0))
plot(0,0,type="n",axes=F,ann=F)
polygon(cos(seq(0,2*pi,length.out=1000)), sin(seq(0,2*pi,length.out=1000)))
for(i in 1:N){
  lines(c(0,cos((i-1)/N*2*pi)),c(0,sin((i-1)/N*2*pi)))
  text(0.75*cos((i-0.5)/N*2*pi),0.75*sin((i-0.5)/N*2*pi),X[i])
}
```

## Example with equally sized wedges

```{r example3,fig.width=2,fig.height=2}
X = c(10,12,14,15,16,18,20)
N = length(X)
par(mar=c(0,0,0,0))
plot(0,0,type="n",axes=F,ann=F)
polygon(cos(seq(0,2*pi,length.out=1000)), sin(seq(0,2*pi,length.out=1000)))
for(i in 1:N){
  lines(c(0,cos((i-1)/N*2*pi)),c(0,sin((i-1)/N*2*pi)))
  text(0.75*cos((i-0.5)/N*2*pi),0.75*sin((i-0.5)/N*2*pi),X[i])
}
arrows(0.5*cos(pi+0.1),0.5*sin(pi+0.1),0.5*cos(0.1),0.5*sin(0.1),length=0.1,code=2,lwd=2)
```
$$\mu = \frac{\sum X}{N} = `r mean(X)`$$
$$\sigma = \sqrt{\frac{\sum(X-\mu)^2}{N}} = `r round(sd(X)*sqrt((N-1)/N),4)` $$


## Running average table

```{r table1}
X = c(10,12,14,15,16,18,20)
N = length(X)
n = 30
x = sample(X,n,T)
```

| $i$ | $x$ | $\sum x$ | $\bar{x} = \frac{\sum x}{i}$ |
|:---:|:---:|:---:|:---:|
| 1 | `r x[1]` | `r sum(x[1])` | `r x[1]` |
| 2 | `r x[2]` | `r sum(x[1:2])` | `r sum(x[1:2])/2` |
| 3 | `r x[3]` | `r sum(x[1:3])` | `r sum(x[1:3])/3` |
| 4 | `r x[4]` | `r sum(x[1:4])` | `r sum(x[1:4])/4` |
| 5 | `r x[5]` | `r sum(x[1:5])` | `r sum(x[1:5])/5` |
| 6 | `r x[6]` | `r sum(x[1:6])` | `r sum(x[1:6])/6` |
| 7 | `r x[7]` | `r sum(x[1:7])` | `r sum(x[1:7])/7` |
| 8 | `r x[8]` | `r sum(x[1:8])` | `r sum(x[1:8])/8` |
| 9 | `r x[9]` | `r sum(x[1:9])` | `r sum(x[1:9])/9` |
| 10 | `r x[10]` | `r sum(x[1:10])` | `r sum(x[1:10])/10` |
| 11 | `r x[11]` | `r sum(x[1:11])` | `r sum(x[1:11])/11` |
|$\vdots$|$\vdots$|$\vdots$|$\vdots$|

## Running-averages plot

```{r example4,fig.width=6,fig.height=4.5}
library(latex2exp)
plot(cumsum(x)/(1:n),type='l',ylim=c(min(X),max(X)),ylab="running mean",xlab=TeX("$i$ (spin count)"),col=rgb(0.2,0.2,0.2),lwd=2)
points(1:n,cumsum(x)/(1:n))
# for(i in 1:8){
#   x = sample(X,n,T)
#   lines(1:n,cumsum(x)/(1:n),col=rgb(0.2,0.2,0.2,0.7),lwd=2)
# }
for(z in -2:2){
  is = 1:n
  mu = mean(X)
  sig = sd(X)*sqrt((N-1)/N)
  lines(is,mu+z*sig/sqrt(is),col=rgb(0.8,0.2,0.2,0.5))
}
```

The smooth curves represent $\mu \pm 2\frac{\sigma}{\sqrt{i}}$ and $\mu \pm \frac{\sigma}{\sqrt{i}}$ and $\mu$. You do not need to draw them, but you can.

## Overlay of multiple running-averages

```{r overlay,fig.width=6,fig.height=4.5}
library(latex2exp)
plot(cumsum(x)/(1:n),type='l',ylim=c(min(X),max(X)),ylab="running mean",xlab=TeX("$i$ (spin count)"),col=rgb(0.2,0.2,0.2,0.7),lwd=2)
for(i in 1:9){
  x = sample(X,n,T)
  lines(1:n,cumsum(x)/(1:n),col=rgb(0.2,0.2,0.2,0.7),lwd=2)
}
for(z in -2:2){
  is = 1:n
  mu = mean(X)
  sig = sd(X)*sqrt((N-1)/N)
  lines(is,mu+z*sig/sqrt(is),col=rgb(0.8,0.2,0.2,0.5))
}
```

This is a **lot** of work by hand. Feel free to overlay multiple running averages if you use a computer. You can just show work for one.

## Sample from $X+X+X$

```{r sum_of_xs}
x1 = sample(X,10,T)
x2 = sample(X,10,T)
x3 = sample(X,10,T)
xsum = x1+x2+x3
```
Each measurement takes three spins.

| $x_1$ | $x_2$ | $x_3$ | $x_1+x_2+x_3$ |
|:---:|:---:|:---:|:---:|
| `r x1[1]` | `r x2[1]` | `r x3[1]` | `r xsum[1]` |
| `r x1[2]` | `r x2[2]` | `r x3[2]` | `r xsum[2]` |
| `r x1[3]` | `r x2[3]` | `r x3[3]` | `r xsum[3]` |
| `r x1[4]` | `r x2[4]` | `r x3[4]` | `r xsum[4]` |
| `r x1[5]` | `r x2[5]` | `r x3[5]` | `r xsum[5]` |
| `r x1[6]` | `r x2[6]` | `r x3[6]` | `r xsum[6]` |
| `r x1[7]` | `r x2[7]` | `r x3[7]` | `r xsum[7]` |
| `r x1[8]` | `r x2[8]` | `r x3[8]` | `r xsum[8]` |
| `r x1[9]` | `r x2[9]` | `r x3[9]` | `r xsum[9]` |
| `r x1[10]` | `r x2[10]` | `r x3[10]` | `r xsum[10]` |
| ===== | ===== | ===== | ===== |
|  |  |  | $\text{total} = `r sum(xsum)`$ |
|  |  |  | $\text{mean} = `r mean(xsum)`$ |
|  |  |  | $\text{stdev} = `r sd(xsum)`$ |

## Sample from $3X$

```{r triple_of_x}
x = sample(X,10,T)
tx = 3*x 
```
Each measurement takes three spins.

| $x$ | $3x$ |
|:---:|:---:|
| `r x[1]` |`r tx[1]` |
| `r x[2]` |`r tx[2]` |
| `r x[3]` |`r tx[3]` |
| `r x[4]` |`r tx[4]` |
| `r x[5]` |`r tx[5]` |
| `r x[6]` |`r tx[6]` |
| `r x[7]` |`r tx[7]` |
| `r x[8]` |`r tx[8]` |
| `r x[9]` |`r tx[9]` |
| `r x[10]` |`r tx[10]` |
| ===== | ===== | 
|  | $\text{total} = `r sum(tx)`$ |
|  | $\text{mean} = `r mean(tx)`$ |
|  | $\text{stdev} = `r sd(tx)`$ |

## Theory (linear combination of random variables)

* If $X$ and $Y$ represent two random variables, and $a$ and $b$ represent two constants, then:
$$SD(aX+bY) = \sqrt{a^2 SD(X)^2+b^2 SD(Y)^2} $$
$$SD(X+Y) = \sqrt{SD(X)^2+SD(Y)^2} $$
$$SD(X+X) = \sqrt{SD(X)^2+SD(X)^2} = \sqrt{2SD(X)^2}= \sqrt{2}SD(X)$$
$$SD(X+X+X) = \sqrt{SD(X)^2+SD(X)^2+SD(X)^2}$$
$$SD(aX) = \sqrt{a^2 SD(X)^2} = a \cdot SD(X)$$
