---
title: "Stats Fundamentals"
author: "Chad Worley"
date: "January 23, 2020"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data (sample) generation

* In the real world, data are generated by randomly sampling from a population.
* In this class, most data are generated with random number generators.
    - Coins
    - Dice
    - Spinners
    - Computer-based random number generators.
* Unfair coins are useful for generating binary data.
* Dice are useful for generating discrete uniform data.
* Spinners and computers can produce any type of data.

## Example data and possible notations

I rolled a 4-sided die 7 times.
```{r dicerolls}
set.seed(123)
sam = sample(1:4,7,T)
```
The results:
`r sam`

We let $x_i$ represent the result of the $i$th roll. We call $i$ the index.
$$\begin{aligned}
x_1 &= `r sam[1]` &  x_2 &= `r sam[2]` &  x_3 &= `r sam[3]` & x_4 &= `r sam[4]` &
x_5 &= `r sam[5]` &  x_6 &= `r sam[6]` &  x_7 &= `r sam[7]` 
\end{aligned}$$
We usually use a table.

| $i$ | $x_i$ |
|:---:|:---:|
|1|`r sam[1]`|
|2|`r sam[2]`|
|3|`r sam[3]`|
|4|`r sam[4]`|
|5|`r sam[5]`|
|6|`r sam[6]`|
|7|`r sam[7]`|

## The summation operator

The summation operator is defined with the following equation.

$$\sum_{i=a}^bx_i ~=~ x_a+x_{a+1}+\dots+x_{b-1}+x_b $$
Simple examples:
$$\sum_{i=1}^{7}x_i ~~=~~ x_1+x_2+x_3+x_4+x_5+x_6+x_7 $$
$$\sum_{i=1}^{3}\sqrt{x_i+6} ~~=~~ \sqrt{x_1+6}+\sqrt{x_2+6}+\sqrt{x_3+6} $$


## The mean

We will always use $n$ for the sample size (number of measurements).
$$n = \text{sample size} $$
The mean is the sum of the values divided by the sample size.

Formal definition:
$$\text{mean} = \cfrac{\sum\limits_{i=1}^n x_i}{n} $$
In this class we can be a bit informal. We always sum over all values implicitly, so we can use a less decorated formula.
$$\text{mean} = \frac{\sum x}{n} $$
We will usually use $\bar{x}$ for a sample mean. We call this "x bar".
$$\bar{x} = \frac{\sum x}{n} $$

## The mean: example

| $i$ | $x_i$ |
|:---:|:---:|
|1|`r sam[1]`|
|2|`r sam[2]`|
|3|`r sam[3]`|
|4|`r sam[4]`|
|5|`r sam[5]`|
|6|`r sam[6]`|
|7|`r sam[7]`|

We calculate the mean by adding all the $x$ values and dividing by the number of values.

$$\bar{x} = \frac{\sum x}{n} = \frac{`r sam[1]`+`r sam[2]`+`r sam[3]`+`r sam[4]`+`r sam[5]`+`r sam[6]`+`r sam[7]`}{7} = `r mean(sam)` $$

## Data = Results = Sample = Measurements = Values

We are going to use a variety of words to mean basically the same thing in this class.

In the past, I have tried using only "measurements", but it feels unnatural. Like when rolling a die, it feels weird to call the results measurements. But, when discussing lizard lengths, it feels weird to call the measurements results.

In the context of rolling dice I will probably say "rolls".

In the context of flipping a coin, I will say "flips".

In the context of spinning a spinner, I will say "spins".

Etc...

## 1-dimensional scatterplots (stripcharts)

Our textbook calls the following a dot plot:

```{r makedata,fig.height=2,fig.width=5}
par(xpd=T)
x = round(runif(20,30,70),2)
xbar = mean(x)
stripchart(x,pch=16,cex = 1.5,col=rgb(0,0,0,0.5))
polygon(c(xbar,xbar+1,xbar-1),c(0.5,0,0))
```

(made from the data below)

`r x`

The triangle marks the mean. Notice it marks the balance point (center of mass).


## dot plots

```{r dotplots,fig.height=2,fig.width=5}
par(xpd=T)
xbar = mean(x)
stripchart(x,pch=16,cex = 1.5,col=rgb(0,0,0,0.5))
polygon(c(xbar,xbar+1,xbar-1),c(0.5,0,0))
```

When we say **dot plot**, we will refer to what the textbook calls a stacked dotplot:

```{r stacked,fig.height=3,fig.width=5}
par(xpd=T)
xbar = mean(x)
stripchart(round(x/5)*5,method="stack",offset=0.5,at=0.2,pch=16,cex = 1.5,col=rgb(0,0,0))
polygon(c(xbar,xbar+1,xbar-1),c(-0.1,-0.3,-0.3))
```

Notice, this was made by first rounding to the nearest multiple of 5. Stacking requires discrete (granular) data. 


## Histograms

A histogram is like a stacked dot plot, but uses bars instead of dots. **We will use histograms a lot.** 

```{r histy,fig.height = 5}
hist(x,breaks=seq(27.5,72.5,by=5))
```

## Center and Spread

We often want to summarize data with center and spread.

* Measures of center:
    - **Mean**
    - Median
* Measures of spread:
    - Range
    - Inter-quartile range
    - Mean absolute deviation
    - **Standard deviation**

The center indicates where the pile is located. The spread indicates how "wide" the pile is, or how much variation the variable has.

## Shapes of histograms

The main shapes I want you to know are: bell-shaped, uniform, bimodal, left-skewed, and right-skewed.

* Bell-shaped describes symmetric mounds with two tails. 

```{r, fig.height=4,fig.width=5,fig.align='center'}
belldata = rbeta(1000,3,3) 
hist(belldata,axes=F,ann=F)
```

## Shapes of histograms

The main shapes I want you to know are: bell-shaped, uniform, bimodal, left-skewed, and right-skewed.

* Uniform describes symmetric mesas (flat-topped hill with steep sides). 

```{r, fig.height=4,fig.width=5,fig.align='center'}
belldata = rbeta(1000,1,1) 
hist(belldata,axes=F,ann=F)
```

## Shapes of histograms

The main shapes I want you to know are: bell-shaped, uniform, bimodal, left-skewed, and right-skewed.

* Bimodal describes two hills around a valley. 

```{r, fig.height=4,fig.width=5,fig.align='center'}
belldata = rbeta(1000,0.5,0.5) 
hist(belldata,axes=F,ann=F)
```

## Shapes of histograms

The main shapes I want you to know are: bell-shaped, uniform, bimodal, left-skewed, and right-skewed.

* Left-skewed describes a mound with a long left tail. 

```{r, fig.height=4,fig.width=5,fig.align='center'}
belldata = rbeta(1000,3,1.3) 
hist(belldata,axes=F,ann=F)
```

## Shapes of histograms

The main shapes I want you to know are: bell-shaped, uniform, bimodal, left-skewed, and right-skewed.

* Right-skewed describes a mound with a long right tail. 

```{r, fig.height=4,fig.width=5,fig.align='center'}
belldata = rbeta(1000,1.6,5) 
hist(belldata,axes=F,ann=F)
```

## The sample standard deviation

The sample standard deviation will always be denoted with $s$.

The standard deviation of a sample (data) is determined with the following formula.

$$s = \sqrt{\frac{\sum\limits_{i=1}^n(x_i-\bar{x})^2}{n-1}} $$
Or, informally:
$$s = \sqrt{\frac{\sum(x-\bar{x})^2}{n-1}} $$

## The sample standard deviation: example calculation

\small
Let's say we have the following data:

6, 9, 11, 8, 12

First, we determine the sample size by counting how many numbers are in the sample.
$$n = 5$$
We can calculate the mean.
$$\bar{x} = \frac{6+9+11+8+12}{5} = 9.2 $$

We can use a table to determine the standard deviation.

| $x$ | $x-\bar{x}$ | $(x-\bar{x})^2$ |  
|:---:|:---:|:---:|
| 6 | -3.2 | 10.24 |
| 9 | -0.2 | 0.04  |
|11 | 1.8 | 3.24   |
|8  | -1.2| 1.44   |
|12 | 2.8 | 7.84   |

## The sample standard deviation: example calculation...

$$ n = 5$$
$$ \bar{x} = 9.2$$

| $x$ | $x-\bar{x}$ | $(x-\bar{x})^2$ |  
|:---:|:---:|:---:|
| 6 | -3.2 | 10.24 |
| 9 | -0.2 | 0.04  |
|11 | 1.8 | 3.24   |
|8  | -1.2| 1.44   |
|12 | 2.8 | 7.84   |

$$\sum(x-\bar{x})^2 = 22.8 $$

$$s = \sqrt{\frac{\sum(x-\bar{x})^2}{n-1}} = \sqrt{\frac{22.8}{5-1}} = \sqrt{5.7} \approx 2.387 $$





